---
title: "CelltypeR Workflow"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Included in this tutorial

1. Data preprocessing
a. Read FlowJo files into R.
b. Create a flow frame with intensity measurements for each marker for all          samples within the experiment to be analyzed.  
c. Harmonize data 
d. Create dataframe with intensity of expression (area) for each marker
e. Create a Seurat object

2. Cluster data
a. Explore clustering methods and parameters.
b. Determine cluster stability
c. Create clusters with best conditions

3. Annotate clusters with cell types.
a. Predict cell types by correlation matrix
b. Visualize expression by cluster for manual annotation
c. With annotated an data set train Random Forest classifier to predict cell        types. (will only work if you have an already labelled dataset and train the model)
d. annotated an data set predict cell types with Seurat label transfer. (will only work if you have a reference labelled data object).
e. For each prediction calculate the top most predicted cell per cluster and take a consensus to annotate clusters. Make a dataframe of predicted annotations. 
f. Add cluster annotations.

4. Compare groups
a. bar charts of proportions
b. permutations tests
c. ANOVAs and tukey tests

5.Training of random forest model.

# Set up
Install CelltypeR and load the library

```{r}
# devtools are required


```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
