---
title: "Figure 4"
output: github_document
---

Figure 4: Visualization of annotated data from the full data set of 9 hMOs.
Annotations methods:
1. Visualization
2. Correlation
3. 
UMAP of cell type annotations
UMAP split by iPSC line (genetic background)

 We begin with the already clustered object.The correlation predictions were also run separately but follow the same code as the workflow. 


```{r}
require(Seurat)
require(tidyverse)
require(CelltypeR)

```
#Annotation of clusters

Read in the seurat object. The data was prepared following the CelltypeR workflow but run in HPC. 

```{r}

seu.q <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/Figure4/All9MOannaoteAug.RDS")

AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")


```


Visualize marker expression
```{r}
Idents(seu.q)<- 'RNA_snn_res.1.5'
for (i in AB) {
  print(FeaturePlot(seu.q, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE, raster = FALSE))
}

```


Heatmap of marker expression by cluster
```{r}

plotmean(plot_type = 'heatmap',seu = seu.q, group = 'RNA_snn_res.1.5',
         markers = AB, 
               var_names = c(0:35), slot = 'scale.data', xlab = "Cluster",
               ylab = "Markers")



```



Visualize Predict cell types with correlation
Correlation predictions were run as in the CelltypeR in HPC.

```{r}
# double cell threshold 0.005
plot_corr(cor, threshold = 0.35, min_cells = 200)

```
```{r, fig.width=6}
# double cell threshold 0.05
cor1 <- read.csv("MO_AllcellsCor_thesh35.csv")
plot_corr(cor1, threshold = 0.35, min_cells = 500)

```

Visualize correlations per cluster by Barchart

```{r, fig.width=5}

Idents(seu.q) <- "RNA_snn_res.1.5"
seu.q <- AddMetaData(seu.q, metadata = seu.q$RNA_snn_res.1.5, 
                     col.name = 'clusters')
seu.q$clusters <- factor(seu.q$clusters, levels = c(0:35))

cor.ann.035 <- get_annotation(seu.q, seu.cluster = seu.q$RNA_snn_res.1.5, 
                          seu.label = seu.q$cor.labels1, top_n = 3, 
                          filter_out = c("Unknown","unknown","Mixed", 
                                         "unassigned","Unassigned"), 
                          Label = "CAM")

# top 5 per groups
t.lables <- as.data.frame(table(seu.q$RNA_snn_res.1.5, seu.q$cor.labels1))
  t.lables$Freq <- as.double(t.lables$Freq)
  colnames(t.lables) <- c("Cluster", "Label","Freq")
  top.labs <- t.lables  %>% group_by(Cluster) %>% top_n(5, Freq)
  sort.tops <- top.labs %>% as.data.frame() %>% arrange(desc(Freq))  %>% arrange(Cluster)
  print(sort.tops)
  
    # make a bar chart of the cell type per cluster
plot_lab_clust(seu.q, seu.q$clusters, seu.q$cor.labels1, filter_out = "unassigned")

# save the plot
pdf("BarchartCorbyClusterThresh035.pdf", width = 15, height = 5)
plot_lab_clust(seu.q, seu.q$clusters, seu.q$cor.labels1, filter_out = "unassigned")
dev.off()
 
 
  

```

Visualize correlation predictions on UMAP per cell, filtering out cell types with less than 500 cells. 

```{r}
Idents(seu.q) <- 'cor.labels1'
# filter cell types less than 500 cells and filter out unassigned from the 0.35 threshold with 0.005 double cell threshold
thresh = 500
freq.cor <- data.frame(table(seu.q$cor.labels1))

keep <- freq.cor %>% filter(Freq > thresh)
# sort by frequency
# sorted_df <- df %>% arrange(desc(Freq))
# alphabetical
sorted_df <- keep %>% arrange(Var1)
cell.order <- as.character(keep$Var1)
cell.order <- cell.order[1:25]

#filter.out <- c(as.character(df.ft$Var1),"unassigned")
seu.sub <- subset(seu.q, ident = cell.order, invert = FALSE)


DimPlot(seu.sub, order = rev(cell.order), raster = FALSE)


# save plots 
png("AllcellsUMAPCorlablesfilteredLB.png", width = 1100, height = 400)
DimPlot(seu.sub, order = rev(cell.order), raster = FALSE, group.by = "cor.labels1", label = TRUE, repel = TRUE, label.size = 5.5)+ theme_classic(base_size = 20)
dev.off()


```

Filter again but keep unassigned

```{r}
Idents(seu.q) <- 'cor.labels1'
# filter cell types less than 500 cells and filter out unassigned from the 0.35 threshold with 0.005 double cell threshold
thresh = 500
freq.cor <- data.frame(table(seu.q$cor.labels1))

keep <- freq.cor %>% filter(Freq > thresh)
# sort by frequency
# sorted_df <- df %>% arrange(desc(Freq))
# alphabetical
sorted_df <- keep %>% arrange(Var1)
cell.order <- as.character(keep$Var1)

#filter.out <- c(as.character(df.ft$Var1),"unassigned")
seu.sub <- subset(seu.q, ident = cell.order, invert = FALSE)

DimPlot(seu.sub, order = rev(cell.order), raster = FALSE)

#png("AllcellsUMAPCorlablesfilteredLB2.png", width = 950, height = 400)
#DimPlot(seu.sub, order = rev(cell.order), raster = FALSE, group.by = "cor.labels1", label = TRUE, repel = TRUE, label.size = 5.5)+ theme_classic(base_size = 20)
#dev.off()


# pdf("BarchartCorbyClusterThresh035Filter2.pdf", width = 9.5, height = 4)
# plot_lab_clust(seu.sub, seu.sub$clusters, seu.sub$cor.labels1, filter_out = "unassigned")
# dev.off()
 
 plot_lab_clust(seu.sub, seu.sub$clusters, seu.sub$cor.labels1, filter_out = "unassigned")

```

Get a table of the top predictions
```{r}
 t.lables <- as.data.frame(table(seu.q$RNA_snn_res.1.5, seu.q$cor.labels1))
  t.lables$Freq <- as.double(t.lables$Freq)
  colnames(t.lables) <- c("Cluster", "Label","Freq")
  top.labs <- t.lables  %>% group_by(Cluster) %>% top_n(5, Freq)
  sort.tops <- top.labs %>% as.data.frame() %>% arrange(desc(Freq))  %>% arrange(Cluster)
  print(sort.tops)


write.csv(sort.tops,"CAMtop5pred.csv")
```

Add the annotations from the correlation prediction


```{r}
seu.q <- CelltypeR::annotate(seu.q,annotations = cor.ann.035$CAM, to_label= "RNA_snn_res.1.5", annotation_name = "CAM")

DimPlot(seu.q, group.by = "CAM", raster = FALSE, label = TRUE)
```


Random Forest Classifier

```{r}

# Random Forest classifier trained with annotations from subset of 9 MBO cells. 
# read in the saved model
rf <- readRDS("/pathway/to/trained/RFM/Seu9000trainedSeptRFM.RDS")

# test random forest function takes in the seurat object to have cells predicted and the trained model.  Must be trained with the same antibodies

rfm.pred <- RFM_predict(seu.q, rf)


# add the predictions to the seurat object
seu.q <- AddMetaData(object=seu.q, metadata= as.factor(rfm.pred$Prediction), col.name = 'rfm.pred')
unique(seu.q$rfm.pred)
Idents(seu.q) <- "rfm.pred"
levels(seu.q)

celltypes <- unique(seu.q$rfm.pred)
sorted <- as.character(factor(sort(as.character(celltypes))))


# change the order of the predicted cells
Idents(seu.q) <- 'rfm.pred'
seu.q <- AddMetaData(seu.q, metadata = seu.q$rfm.pred, col.name = 'rfm.sort')
seu.q$rfm.sort <- factor(seu.q$rfm.sort, levels = sorted)


# save plots
 pdf("BarchartRFMbyClusterSept.pdf", width = 8, height = 4)
 plot_lab_clust(seu.q, seu.q$clusters, seu.q$rfm.sort, filter_out = c( "Unassigned","Mixed"))
 dev.off()
 
png("UMAPRFMpredictions.png", width = 800, height = 400)
DimPlot(seu.q, group.by = "rfm.pred", raster = FALSE,  label = TRUE, 
        repel = TRUE, label.size = 5.5,
        order = rev(sorted)) + theme_classic(base_size = 20)
 dev.off()
 
 
 # plot UMAP and bar chart
DimPlot(seu.q, group.by = 'rfm.pred', raster = FALSE, label = TRUE, order = rev(sorted))
plot_lab_clust(seu.q, seu.q$clusters, seu.q$rfm.sort, filter_out = c( "Unassigned","Mixed"))

rfm.ann <- get_annotation(seu.q, seu.cluster = seu.q$RNA_snn_res.1.5, 
                          seu.label = seu.q$rfm.pred, top_n = 3, 
                          filter_out = c("Mixed","Mix", 
                                         "unassigned","Unassigned"), 
                          Label = "RFM")

 
 

```

```{r}
 
 t.lables <- as.data.frame(table(seu.q$RNA_snn_res.1.5, seu.q$rfm.pred))
  t.lables$Freq <- as.double(t.lables$Freq)
  colnames(t.lables) <- c("Cluster", "Label","Freq")
  top.labs <- t.lables  %>% group_by(Cluster) %>% top_n(5, Freq)
  sort.tops <- top.labs %>% as.data.frame() %>% arrange(desc(Freq))  %>% arrange(Cluster)
  print(sort.tops)

write.csv(sort.tops,"RFMtop5predNew.csv")
```
Add annotation from the RFM predictions

```{r}
seu.q <- annotate(seu.q, annotations = rfm.ann$RFM, to_label= "RNA_snn_res.1.5", annotation_name = "RFM")

DimPlot(seu.q, group.by = "RFM", raster = FALSE, label = TRUE)
```


Seurat label transfer

```{r}
seu.r <- readRDS("pathway/to/saved/reference/LabelledSeurat.RDS")
# notes this function automatically selects the total number of markers as the dim.  So dim = 1:13 here. 
seu.q <- seurat_predict(seu.q, seu.r, ref_id = 'labels', 
                        refdata = seu.r$labels,
                           down.sample = 500, 
                        markers)

DimPlot(seu.q, group.by = 'seu.pred', raster = FALSE, label = TRUE) + theme_classic(base_size = 15)
setwd("~/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/Figure4")
 png("UMAP_seurat_predictionsLB.png", width = 800, height = 400)
DimPlot(seu.q, group.by = 'seu.pred', raster = FALSE, label = TRUE, 
        repel = TRUE, label.size = 5.5) + theme_classic(base_size = 20)
 dev.off()

 

```


Barchart of the  number of cell predicted in each cluster

```{r}
plot_lab_clust(seu.q, seu.q$clusters, seu.q$seu.pred, filter_out = "Unassigned")

#save plot
 pdf("BarchartSeuratbyCluster.pdf", width = 8, height = 4)
 plot_lab_clust(seu.q, seu.q$clusters, seu.q$seu.pred, filter_out = "Unassigned")

```

```{r}
 t.lables <- as.data.frame(table(seu.q$RNA_snn_res.1.5, seu.q$seu.pred))
  t.lables$Freq <- as.double(t.lables$Freq)
  colnames(t.lables) <- c("Cluster", "Label","Freq")
  top.labs <- t.lables  %>% group_by(Cluster) %>% top_n(5, Freq)
  sort.tops <- top.labs %>% as.data.frame() %>% arrange(desc(Freq))  %>% arrange(Cluster)
  print(sort.tops)

write.csv(sort.tops,"Seutop5pred.csv")
```

See the predictions for each method 

```{r}

 seu.ann <- get_annotation(seu.q, seu.cluster = seu.q$RNA_snn_res.1.5, 
                          seu.label = seu.q$seu.pred, top_n = 3, 
                          filter_out = c("Mixed","Mix", 
                                         "unassigned","Unassigned"), 
                          Label = "SEU")

# all the dataframes need to be as.factor
ann.list <- list(cor.ann.035,rfm.ann,seu.ann)

# get a table of the most freq predicted cell by cluster
an.df <- annotate_df(ann.list)

setwd("~/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/Figure4") 
write.csv(an.df,"TopPredCon.csv")

```
Add annotations

```{r, fig.width=4}

cluster.ids <- c("Glia-lineage","Radial Glia 1a","Radial Glia 1",
                 "Radial Glia 3","Neurons 1", #4
                 "Radial Glia 2","Epithelial","Radial Glia 1",
                 "Neurons 1","Radial Glia 1",    #9
                 "Astrocytes 2","Neural-lineage","Endothelial",
                 "Radial Glia 1a","Neurons 1", #14
                 "Astrocytes 1","NPC","Neurons 2","Neurons 2",
                 "Astrocytes 1", #19
                 "Neurons 2","Astrocytes 1","OPC-like","Astrocytes 1","OPC-like", #24
                 "Radial Glia 1","Astrocytes 1","Neurons 2","Oligodendrocytes","OPC",
                 "Radial Glia 2","NPC","Astrocytes 2","Endothelial",
                 "Stem cell like","Radial Glia 1")

# note Neurons 3 is NPC


seu.q <- annotate(seu.q, annotations = cluster.ids, 
                  to_label= "clusters", 
                  annotation_name = "Annotations")

DimPlot(seu.q,label = TRUE, repel = TRUE ,raster = FALSE,
        label.size = 6)



```




```{r, fig.width=4}


# order the cell types
cell.order <- c("Astrocytes 1", "Astrocytes 2",
                     "Radial Glia 1","Radial Glia 1a","Radial Glia 2",
                     "Radial Glia 3","Glia-lineage",
                     "Epithelial","Endothelial",
                     "Neurons 1","Neurons 2","NPC","Neural-lineage",
                     "Oligodendrocytes","OPC","OPC-like",
                     "Stem cell like")



# colour order to match cell type order
clust.colours <- c("chocolate2","darkorange",
                   "pink","deeppink","plum1","lightpink3",
                   "mistyrose2",
                   "steelblue3","deepskyblue",
                   "mediumpurple1","purple","plum3","mediumslateblue",
                   "seagreen3","olivedrab4","darkseagreen3",
                   "tomato3","burlywood3")
 

setwd("~/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/Figure4")

pdf("UMAP.all.groups.pdf",width = 9, height = 4.5)
DimPlot(seu.q, cols = clust.colours, shuffle = TRUE, 
        raster=FALSE, pt.size = 0.1, label = FALSE, 
        group.by = 'Annotations',
        order = rev(cell.order)) +
theme(legend.text = element_text(size=16), axis.title.y = element_text(size=16), 
         axis.title.x = element_text(size=16), axis.text.y = element_text(size =16),
        axis.text.x = element_text(size =16))
dev.off()


DimPlot(seu.q, cols = clust.colours, shuffle = TRUE, 
        raster=FALSE, pt.size = 0.1, label = TRUE, 
        group.by = 'Annotations',
        order = rev(cell.order))


```


Make a dot plot of expression in full dataset


```{r, fig.width=2}
# Figure 4B
#library(data.table)
#Idents(seu.q) <- "Annotations"
#annotations <- levels(seu.q)



dp <- plotmean(plot_type = 'dotplot',seu = seu.q, group = 'Annotations',
         markers = AB, 
               var_names = annotations, slot = 'scale.data', xlab = "Cell type",
               ylab = "Markers",
         cluster_order=rev(cell.order), marker_order=NULL,low_colour = "#154c79",mid_colour = "#eeeee4",
                     high_colour = "#e28743")


# save 
setwd("~/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/Figure4")

pdf("DotPlotCelltypes.pdf",width = 8, height = 5)
print(dp)
dev.off()

dp



```

To compare between variables we need to add the metadata into the seurat object. 
The sample names have the required information

```{r}
# all the data you have should correspond to different samples
# be sure to check the name in the meta data
colnames(seu.q@meta.data)
# when added metadata, like naming the clusters you need to know the order of the levels. 
# we can see this by setting the ident to the metadata we want to name. 
Idents(seu.q) <- 'Sample'
# now print out the levels, this is the order they appear in the data object
levels(seu.q)

# we need to enter a vector in the order we want the meta data 
# this could be read in from a spreadsheet but then we wouldn't see the information here in the workbook. 
# a csv file could be made with the metadata name as the columns, in the order of the levels. Then read in the csv as a data frame and the ipsc.line <- df$ipsc
# the same can be done for each metadata

# we can use the annotate function to add all the metadata
ipsc.lines <- c("3450","3450","3450","AIW002","AIW002","AIW002","AJG001","AJG001","AJG001")
seu.q <- annotate(seu.q, annotations = ipsc.lines, 
                  to_label= "Sample", 
                  annotation_name = "IPSC")

# add experiment date
ex.date <- c("0306","0317","0317","0306","0317","0317","0306","0317","0317")
seu.q <- annotate(seu.q, annotations = ex.date, 
                  to_label= "Sample", 
                  annotation_name = "ExDate")

# add batch information
batchnum <- c("B","A","B","B","A","B","B","A","B")
seu.q <- annotate(seu.q, annotations = batchnum, 
                  to_label= "Sample", 
                  annotation_name = "Batch")

# add organoid days in final culture
days <- c("273","263","284","273","263","284","273","263","284")
seu.q <- annotate(seu.q, annotations = days, 
                  to_label= "Sample", 
                  annotation_name = "Days")
# now you can see that new metadata is added
colnames(seu.q@meta.data)

# to check if each is correct use 
unique(seu.q$Batch)  # change the meta data to each in turn.

```


Run split UMAP - split by iPSC line

```{r, fig.width=5}

# cell order from above
# colours from above

# to change the order of the plots for the split by 
# to do that we need to factor the levels
seu.q$IPSC <- factor(seu.q$IPSC, levels = c("AIW002","AJG001","3450"))



setwd("~/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/Figure4")

pdf("UMAPsplitbyIPSCline.pdf",width = 20, height = 4.5)
DimPlot(seu.q, cols = clust.colours, shuffle = TRUE, 
        raster=FALSE, pt.size = 0.1, label = FALSE, 
        group.by = 'Annotations',
        order = rev(cell.order), split.by = "IPSC") +
theme(legend.text = element_text(size=16), axis.title.y = element_text(size=16), 
         axis.title.x = element_text(size=16), axis.text.y = element_text(size =16),
        axis.text.x = element_text(size =16))
dev.off()

DimPlot(seu.q, cols = clust.colours, shuffle = TRUE, 
        raster=FALSE, pt.size = 0.1, label = FALSE, 
        group.by = 'Annotations',
        order = rev(cell.order), split.by = "IPSC")

```


Plot the proportions of cell types

```{r}
# to get the correct order for the plot again we need to refactor
Idents(seu.q) <- "Annotations"
seu.q$Celltypes <- seu.q$Annotations


# cell order and clust.colours defined above)
seu.q$Celltypes <- factor(seu.q$Celltypes, levels = c("Astrocytes 1", "Astrocytes 2",
                     "Radial Glia 1","Radial Glia 1a","Radial Glia 2",
                     "Radial Glia 3","Glia-lineage",
                     "Epithelial","Endothelial",
                     "Neurons 1","Neurons 2","NPC","Neural-lineage",
                     "Oligodendrocytes","OPC","OPC-like",
                     "Stem cell like"))


proportionplots(seu.q,seu.var = seu.q$IPSC, seu.lable = seu.q$Celltypes, groups = "IPSC", my_colours = clust.colours)

setwd("~/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/Figure4")
pdf("PercentCelltypesIPSC.pdf", width = 6, height = 4)
proportionplots(seu.q,seu.var = seu.q$IPSC, seu.lable = seu.q$Celltypes, groups = "IPSC", my_colours = clust.colours)
dev.off()

```


We can look at the other variables

```{r}
proportionplots(seu.q,seu.var = seu.q$Days, seu.lable = seu.q$Celltypes, groups = "Days in culture", my_colours = clust.colours)

proportionplots(seu.q,seu.var = seu.q$Batch, seu.lable = seu.q$Celltypes, groups = "hMO Batch", my_colours = clust.colours)

proportionplots(seu.q,seu.var = seu.q$Ex.date, seu.lable = seu.q$Celltypes, groups = "Experiment Date", my_colours = clust.colours)

```


Proportion tests

```{r}

#library(scProportionTest)
# create a propotion test object

Idents(seu.q) <- "Celltypes"
prop_test <- sc_utils(seu.q)

# set up the comparison

# need to compare separately
prop_test <- permutation_test(
	prop_test, cluster_identity = "Celltypes",
	sample_1 = "AIW002", sample_2 = "AJG001",
	sample_identity = "IPSC"
)

# make the plot
permutation_plot(prop_test)


# save the plot 
png("prp.test.AIWvAJG.png", width = 600, height = 300)
permutation_plot(prop_test) +  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),    # Adjust axis title size
  legend.text = element_text(size = 12),   # Adjust legend text size
  legend.title = element_text(size = 14))
dev.off()



 

setwd("~/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/Figure4")

prop_test <- permutation_test(
	prop_test, cluster_identity = "Celltypes",
	sample_1 = "AIW002", sample_2 = "3450",
	sample_identity = "IPSC"
)


# make the plot
permutation_plot(prop_test)

# save the plot 
png("prp.test.AIWv3450.png",width = 600, height = 300)
permutation_plot(prop_test) +  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),    # Adjust axis title size
  legend.text = element_text(size = 12),   # Adjust legend text size
  legend.title = element_text(size = 14))
dev.off()


prop_test <- permutation_test(
	prop_test, cluster_identity = "Celltypes",
	sample_1 = "3450", sample_2 = "AJG001",
	sample_identity = "IPSC"
)


# make the plot
permutation_plot(prop_test)

# save the plot 
png("prp.test.3450vAJG.png", width = 600, height = 300)
permutation_plot(prop_test) +  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),    # Adjust axis title size
  legend.text = element_text(size = 12),   # Adjust legend text size
  legend.title = element_text(size = 14))
dev.off()

# make a group of AIW vs other
#Idents(seu) <- "Batch"
#cluster.ids <- c("AJG-3450","AJG-3450","AJG-3450","AIW002","AIW002","AIW002",
#                 "AJG-3450","AJG-3450","AJG-3450")
#names(cluster.ids) <- levels(seu)
#seu <- RenameIdents(seu, cluster.ids)
#seu$ipsc <- Idents(seu)

# join together AJG and 3450
prop_test <- permutation_test(
	prop_test, cluster_identity = "Celltypes",
	sample_1 = "AIW002", sample_2 = "AJG-3450",
	sample_identity = "ipsc")

# make the plot
permutation_plot(prop_test)
# save the plot 
pdf("prp.test.AIWvs3450-AJG.pdf",width = 9, height = 4)
permutation_plot(prop_test) + theme_bw() + 
  theme(axis.text.x=element_text(size=15),
                     axis.text.y=element_text(size=15))
dev.off()

```


Proportionality test for other variables
```{r}
# prop_test prepared above
setwd("~/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/Figure4")
# compare batch
prop_test <- permutation_test(
	prop_test, cluster_identity = "Celltypes",
	sample_1 = "A", sample_2 = "B",
	sample_identity = "Batch"
)


# save the plot 
png("prp.test.AvB_batch.png", width = 600, height = 300)
permutation_plot(prop_test) +  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),    # Adjust axis title size
  legend.text = element_text(size = 12),   # Adjust legend text size
  legend.title = element_text(size = 14))
dev.off()

# Exdate
prop_test <- permutation_test(
	prop_test, cluster_identity = "Celltypes",
	sample_1 = "0306", sample_2 = "0317",
	sample_identity = "ExDate"
)

# save the plot 
png("prp.test.compareDate.png", width = 600, height = 300)
permutation_plot(prop_test) +  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),    # Adjust axis title size
  legend.text = element_text(size = 12),   # Adjust legend text size
  legend.title = element_text(size = 14))
dev.off()



```



Heat map grouped byIPSC line


```{r}
library(ggplot2)
library(dplyr)
library(tidyr)
library(tibble)  # For rownames_to_column function
# Sample data
# Assuming df is your data frame with the structure you described

# Extract just the marker names from the column names
marker_names <- gsub("^RNA\\.", "", colnames(df))

# Create a new data frame with the required structure
heatmap_data <- df %>%
  rownames_to_column(var = "Markers") %>%  # Use rownames_to_column from the tibble package
  pivot_longer(-Markers, names_to = "Celltype_IPSC", values_to = "Expression")

# Separate Celltype and IPSC

# Modify the splitting pattern
heatmap_data <- heatmap_data %>%
  separate(Celltype_IPSC, into = c("Celltype", "IPSC"), sep = "_", remove = FALSE) %>%
  mutate(Celltype = gsub("^RNA\\.", "", Celltype))  # Remove "RNA." from Celltypes



# Specify the order of Celltypes and IPSC
celltype_order <- c("Astrocytes.1", "Astrocytes.2",
                     "Radial.Glia.1","Radial.Glia.1a","Radial.Glia.2",
                     "Radial.Glia.3","Glia.lineage",
                     "Epithelial","Endothelial",
                     "Neurons.1","Neurons.2","NPC","Neural.lineage",
                     "Oligodendrocytes","OPC","OPC.like",
                     "Stem.cell.like")

ipsc_order <- c("AIW002", "AJG001", "3450")  # Define the order of IPSC

# Reorder the levels of Celltypes and IPSC
heatmap_data$Celltypes <- factor(heatmap_data$Celltype, levels = celltype_order)
heatmap_data$IPSC <- factor(heatmap_data$IPSC, levels = ipsc_order)

# Create the heatmap
ggplot(heatmap_data, aes(x = interaction(IPSC, factor(Celltype, 
                                                      levels = celltype_order)), 
                         y = factor(Markers, levels = rev(AB)), fill = Expression)) +
  geom_tile() +
  scale_fill_gradient2(low = "#154c79", mid = "#eeeee4", high = "#e28743") +
  labs(x = "Cell Types and IPSC", y = "Markers") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 12))

```


Save the heatmap grouped by iPSC line and Celltype

```{r}
# heatmap

setwd("~/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/Figure4")
pdf("heatmap.ave.line.cell.pdf", width = 13, height = 4)
ggplot(heatmap_data, aes(x = interaction(IPSC, factor(Celltype, 
                                                      levels = celltype_order)), 
                         y = factor(Markers, levels = rev(AB)), fill = Expression)) +
  geom_tile() +
  scale_fill_gradient2(low = "#154c79", mid = "#eeeee4", high = "#e28743") +
  labs(x = "Cell Types and IPSC", y = "Markers") +
  theme_bw() +
  theme(axis.text.x = element_text(size = 12, angle = 45, hjust = 1, vjust = 1),
        axis.text.y = element_text(size = 12),
        plot.title = element_text(size = 12))
dev.off()

```


Statistic comparing expression between groups.

```{r}

# first we use the function top prepare a data frame that will be the input for the stats function
# list the metadata slot names for desired variables, you need these to match what is in metadata
colnames(seu.q@meta.data)
var.names <- c("Sample","IPSC","Days", "Batch", "ExDate", "Celltypes")
df.for.stats <- Prep_for_stats(seu.q, marker_list = AB, variables = var.names, 
                               marker_name = 'Marker')
# check to see if the correct information is selected
# we expect columns for each metadata in var.names and "value" which is the expression value
# rows are cells and we expect the dim to match the seurat object
head(df.for.stats)
dim(df.for.stats)

# the celltype needs to be named "CellType"

colnames(df.for.stats) <- c("Sample","IPSC","Days","Batch",
                            "ExDate","CellType","Marker","value")


# With one way anova we will compare differences across all marker expression
# here we will run an ANOVA across IPSC lines for each celltype
# post hoc Tukey's HSD test will tell us which cell lines are different
test.stats1 <- run_stats(input_df= df.for.stats, 
                         group_cols = c("Sample","IPSC", "CellType","Marker"),
                     value_col = "value",
                     stat_type = "ANOVA", id1 = 'IPSC', 
                     id2 = NULL, use_means = TRUE,
                     loop_by = "CellType")

# see the dataframe results
anv <- test.stats1[['ANOVA']]
tk <- test.stats1[['TukeyHSD']]

# save the results. 
library(writexl)
setwd("~/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/Figure4")
write_xlsx(test.stats1, path = "ANOVAcompareIPSC.xlsx")



# with use_means = TRUE we are taking the mean per sample which give us an n=3 for each IPSC line
# if use_means = FALSE the cells will be used as n and our degrees of freedom would be very high

head(anv)
head(tk)


test.stats2 <- run_stats(input_df= df.for.stats, 
                         group_cols = c("Sample","IPSC", "CellType","Marker"),
                     value_col = "value",
                     stat_type = "ANOVA", id1 = 'IPSC', 
                     id2 = NULL, use_means = FALSE,
                     loop_by = "CellType")

# see the dataframe results
anv2 <- test.stats2[['ANOVA']]
tk2 <- test.stats2[['TukeyHSD']]

# with use_means = TRUE we are taking the mean per sample which give us an n=3 for each IPSC line
# if use_means = FALSE the cells will be used as n and our degrees of freedom would be very high

head(anv2)
head(tk2)



```


```{r}
test.stats3 <- run_stats(input_df= df.for.stats, 
                         group_cols = c("Sample","IPSC","Days","Batch",
                            "ExDate","CellType"),
                     value_col = "value",
                     stat_type = "ANOVA", id1 = 'Days', 
                     id2 = NULL, use_means = TRUE,
                     loop_by = "CellType")

test.stats3[['ANOVA']]
test.stats3[['TukeyHSD']]

test.stats4 <- run_stats(input_df= df.for.stats, 
                         group_cols = c("Sample","IPSC","Days","Batch",
                            "ExDate","CellType"),
                     value_col = "value",
                     stat_type = "ANOVA", id1 = 'Batch', 
                     id2 = NULL, use_means = TRUE,
                     loop_by = "CellType")

test.stats4[['ANOVA']]
test.stats4[['TukeyHSD']]

test.stats5 <- run_stats(input_df= df.for.stats, 
                         group_cols = c("Sample","IPSC","Days","Batch",
                            "ExDate","CellType"),
                     value_col = "value",
                     stat_type = "ANOVA", id1 = 'ExDate', 
                     id2 = NULL, use_means = TRUE,
                     loop_by = "CellType")

test.stats5[['ANOVA']]
test.stats5[['TukeyHSD']]


```



2 way anova to compare proteins separately

```{r}
test.stats6 <- run_stats(input_df= df.for.stats, group_cols = c("Sample", "CellType","Marker","IPSC","Batch"),
                     value_col = "value",
                     stat_type = "ANOVA2", id1 = 'IPSC', 
                     id2 = "Marker", use_means = TRUE,
                     loop_by = "CellType")


# overall effects
test.stats6[['ANOVA']]

setwd("~/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/Figure4")
write.csv(test.stats6[['ANOVA']], "ANOVA2maineffectsIPSCMarker.csv")

names(test.stats6$TukeyHSD) <- c("IPSC", "Marker","IPSC-Marker",
                                 "Interactions_ IPSC",
                                 "Interactions_ Marker")

write_xlsx(test.stats6$TukeyHSD, "ANOVA2celltypesMarkersIPSC.xlsx")
# see the relevant interactions Tukey's results
df <- as.data.frame(test.stats6$TukeyHSD$`Interactions_ IPSC`)
head(df)

# now filter for significant differences

df_sig <- df %>% filter(p.adj < 0.05)
df_sig


```

If we use all cells as n do we have more significant differences?


Plot Protein expression by iPSC line.

 test new stats function to fix loop by "CellType" requirement
 
```{r}
test.stats <- run_stats2(input_df= df.for.stats, group_cols = c("Sample", "CellType","Marker","IPSC","Batch"),
                     value_col = "value",
                     stat_type = "ANOVA", id1 = 'IPSC', 
                     id2 = NULL, use_means = TRUE,
                     loop_by = "CellType")



```
 test syntax
 
```{r}

df_means <- get_means(df.for.stats,group_cols = c("Sample", "CellType","Marker","IPSC","Batch"), value_col = "value")

loop_by = "CellType"

var_list <- unique(df_means[[loop_by]])

id1 = "IPSC"
one_way <- aov(expression ~ df_means[[id1]], data = df_means)

```
 

