---
title: "Figure 3"
output: github_document
---

# Figure 3
Data processing and plots

# In this workbook
1. Read in a subset of fsc live gated single cells.
2. Align data, create expression data frames.
3. Create Seurat object and cluster. 
4. Visualize expression
5. Correlation predictions
 a) visualize predicted expression reference matrix
 b) visualize correlation predictions.
 c) compare thresholds of correlations.
6. Annotate.
7. Train random forest classifier.

```{r}
# set up the environment

library(Seurat)
library(dplyr)
library(ggplot2)
library(reshape2)
library(CelltypeR)

```

# Process data

Read in the live gated cells in fsc files. 
```{r}
input_path <- "/Users/rhalenathomas/GITHUB/CelltypeR/FlowCytometry_Data/9MBO/"

flowset <- fsc_to_fs(input_path, downsample = 9000)
# down sample can be a number, 'none' or 'min'

# look at file names and rename with shorter sample names
library(flowWorkspace)  # this library has the function "sampleNames"
sampleNames(flowset)

sampleNames(flowset) <- sampleNames(flowset) <- c("3450_0306","AIW002_0306","AJG001C_0306","3450_0317A","AIW002_0317A","AJG001C_0317A","3450_0317B","AIW002_0317B","AJG001C_0317B")
sampleNames(flowset)


```
Plots to find which measures have two peaks for alignement

```{r}
flowset_biexp <- harmonize(flowset, processing = 'biexp')

# we need to view the biex transform to see which measurements have two peaks
plotdensity_flowset(flowset_biexp)
```
Align peaks and retrotransform

```{r}
flowset_retro <- harmonize(flowset, processing = 'retro', 
                           two_peaks = c(7:20),
                       one_peak = c(1:6,21), threshold = 0.01)


# now create the data frame from the flowset object using the CelltypeR function flowset_to_csv
df <- flowset_to_csv(flowset_retro)
write.csv(df, "MOsflowset9000subset_df.csv")

```


Make Seurat object and cluster

```{r}
AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

# create the seurat object
seu <- make_seu(df, AB_vector = AB)

```

Plot with ggplot Figure 3A

```{r}

# prepare the matrix
reference_path <- "/Users/rhalenathomas/GITHUB/CelltypeR/ExampleOuts/ReferenceMatrices/FinalReferenceMatrix.csv"
reference_data <- read.csv(reference_path)
# the reference matrix need to be in the format cell types as rows and markers as columns
# there is a column X with the markers
#colnames(reference_data)
# we need a row X with the cell type names
# Transpose the dataframe and convert values to numeric

rownames(reference_data) <- reference_data$X
reference_data <- reference_data %>% select(-"X")
colnames(reference_data) <- c("Astrocyte","Endothelial","Epithelial","Neuron",
                   "NPC","OPC","Oligo","RG","Stem")


mat <- as.matrix(reference_data)


marker.order <- rev(c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4"))
# reformat df
long_df <- melt(mat, varnames = c("Marker", "Cell_Type"), value.name = "Expression")
long_df$Marker <- factor(long_df$Marker, levels = marker.order)

hm <- ggplot(long_df, aes(x = Cell_Type, y = Marker, fill = Expression)) +
  geom_tile(color = "white") +  # Remove the black border by setting color to "white"
  scale_fill_gradient(low = "seashell2", high = "red3") +
  labs(x = "Cell Type", y = "Marker", fill = "Relative Expression") + 
  theme_bw() +
  theme(panel.border = element_blank(),  # Remove the panel border
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size = 12, colour = "black"), 
        axis.text.y = element_text(hjust = 1 ,vjust = 0.5, size = 12, colour = "black"),
        axis.title = element_text(colour = "black", size = 14),
        panel.grid = element_blank(),
        axis.ticks = element_blank())  # Remove the panel grid

hm

pdf("/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/June2023/ReferenceMatrix.pdf", width = 6.5, height = 5)
hm
dev.off()


```


Calculate correlations

```{r}

test_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/9MBO/prepro_outsjan20-9000cells/prepro_outsretrotransformed_flowset.csv"

test.df <- read.csv(test_path)
dim(test.df)
head(test.df)

# with 13 antibodies R value of 0.553 has a significant p value less than 0.05
cor1 <- find_correlation(test = test.df, reference = df_ref2,
                        min_corr = 0.553, min_diff = 0.05)

write.csv(cor1, "cor9000cellsRthresh55.csv")

# see how the cells mostly would be assigned if not for the threshold
cor2 <- find_correlation(test = test.df, reference = df_ref2,
                        min_corr = 0.1, min_diff = 0.05)

write.csv(cor2, "cor9000cellsRthresh01.csv")


cor3 <- find_correlation(test = test.df, reference = df_ref2,
                        min_corr = 0.35, min_diff = 0.05)

write.csv(cor3, "cor9000cellsRthresh35.csv")


```

Plot correlations
Figure 3B, C and S3

```{r}

# plot the main groups - and the correlation co-efficient for the assigned group
# plots are created by the plot_corr function, which uses ggplot2
plot_corr(cor1, threshold = 0.553, min_cells = 70) 
plot_corr(cor2, threshold = 0.1, min_cells = 500)
plot_corr(cor3, threshold = 0.35, min_cells = 500)

# save the plots
fig_outs <- "/Users/rhalenathomas/Documents/Projects_Papers/PhenoID/ForFigures/June2023/"

p.cor1 <- plot_corr(cor1, threshold = 0.553, min_cells = 70)

pdf(paste(fig_outs,"CorPlots_thresh0553.pdf",sep = ""))
p.cor1
dev.off()

p.cor2 <- plot_corr(cor2, threshold = 0.1, min_cells = 500)



pdf(paste(fig_outs,"CorPlots_thresh01.pdf",sep = ""))
p.cor2
dev.off()
#How many cells are there in each category - this is in the frequency table
freq.cor2.df <- as.data.frame(p.cor2[[1]])

p.cor3 <- plot_corr(cor3, threshold = 0.35, min_cells = 500)

pdf(paste(fig_outs,"CorPlots_thresh035.pdf",sep = ""))
p.cor3
dev.off()
#How many cells are there in each category - this is in the frequency table
freq.cor3.df <- as.data.frame(p.cor3[[1]])



```

Adjust pair of cell types plots to plot mixes with higher cell counts
Figures S4,S5,S6 

```{r}
# Define the threshold number of pairs
df <- cor1
threshold <- 8

# Filter the double-labeled cells
double.cells <- df[grep("-", df$cell.label),]
label_counts <- table(double.cells$cell.label)
filters_labels <- names(label_counts[label_counts >= threshold])
# Filter the double-labeled cells based on the filtered labels
filtered_double.cells <- double.cells[double.cells$cell.label %in% filters_labels, ]

# Melt the filtered double-labeled cells for plotting
df.melt.filtered <- melt(filtered_double.cells)

# Plot the filtered double-labeled cells
p1 <- ggplot(df.melt.filtered, aes(x = variable, y = value, colour = variable, group = X))+
      geom_line(show.legend = FALSE, size = 0.1, color = "black") +
      geom_point() +
      scale_color_manual(values = c("#4E84C4", "#52854C","purple","orange")) +
      ylim(0, 1) +
      facet_wrap(~ as.factor(cell.label)) +
      ylab("Correlation Coefficient") +
      xlab("")

p1

pdf(paste(fig_outs,"CAMcorpairs_thresh0553.pdf",sep=""))
p1
dev.off()

```


```{r}
# Define the threshold number of pairs
df <- cor3 # threshold 0.35
threshold <- 150

# Filter the double-labeled cells
double.cells <- df[grep("-", df$cell.label),]
label_counts <- table(double.cells$cell.label)
filters_labels <- names(label_counts[label_counts >= threshold])
# Filter the double-labeled cells based on the filtered labels
filtered_double.cells <- double.cells[double.cells$cell.label %in% filters_labels, ]

# Melt the filtered double-labeled cells for plotting
df.melt.filtered <- melt(filtered_double.cells)

# Plot the filtered double-labeled cells
p2 <- ggplot(df.melt.filtered, aes(x = variable, y = value, colour = variable, group = X))+
      geom_line(show.legend = FALSE, size = 0.1, color = "black") +
      geom_point() +
      scale_color_manual(values = c("#4E84C4", "#52854C","purple","orange")) +
      ylim(0.1, 0.8) +
      facet_wrap(~ as.factor(cell.label)) +
      ylab("Correlation Coefficient") +
      xlab("")

p2

pdf(paste(fig_outs,"CAMcorpairs_thresh35.pdf",sep=""))
p2
dev.off()

```



Make plots for Figure 3B and C

```{r}

# plotting script from inside the function above
threshold = 0.553
min_cells = 55

# violin plot
df.filter <- cor1 %>% group_by(cell.label) %>% dplyr::filter(n()> min_cells)

plot2 <- ggplot(df.filter, aes(x = best.cell.type, y = cor.1, fill = best.cell.type)) +
    geom_violin(trim = FALSE) +
    ylim(0, 1) +
    theme_classic() +
    theme(text = element_text(size = 14), axis.text.x = element_text(angle = 90, size = 12)) +
    ylab("correlation coefficient") +
    xlab("Cell type with max correlation coefficient") +
    geom_hline(yintercept = threshold) +
    guides(fill = guide_legend(title = "Cell Type"))


# save
pdf(paste(fig_outs,"VlnCAMthresh533.pdf",sep=""),width = 5,height = 4)
plot2
dev.off()
# bar chart with the R > 0.553 threshold

df.filter2 <- cor1 %>%
    filter(!cell.label %in% c("Unassigned", "unassigned")) %>%
    group_by(cell.label) %>%
    filter(n() > min_cells)
  

  plot1b <-
    ggplot(df.filter2, aes(x = reorder(
      cell.label, cell.label, function(x) - length(x)),
      fill = cell.label)) + geom_bar() + theme_classic() +
    theme(axis.text.x = element_text(size = 12, colour = "black", angle = 90, hjust=0.99,vjust=0.5),
          axis.text.y = element_text(size = 12, colour = "black"),
          axis.title.x = element_text(size = 14, colour = "black"),
          axis.title.y = element_text(size = 14, colour = "black"),
          plot.margin = margin(15, 1, 1, 1)) +

    scale_y_continuous(expand = c(0, 0)) + # take the space away below the bars
    xlab('Assigned cell type') +
    ylab('number of cell') +
    labs(fill='Cell Types')
  

pdf(paste(fig_outs,"BarchartCAMthresh533.pdf",sep=""),width = 5,height = 4)
plot1b
dev.off()

plot1b  
plot2

```

Figure 3 clustering

This is with the subsample of 9000 cells per hMO
Figure 3D and E


```{r}
# read in the seurat objects
seu <- get_clusters(seu, method = "louvain",
                         df_input = df,
                         k = 60,
                         pcdim = 1:12,
                         resolution = 0.7,
                         plots = FALSE)

# the highest Rand Index are res = 0.1, 7 clusters, res = 0.15, 9 clusters.  Both very low RI std.
# the high RI with low std is 0.3 and 0.7, cluster numbers also have low std 
# from bootstrap 100X
# annotation is easier with 18-25 clusters 

seu <- RunUMAP(seu, spread = 1, min.dist = 0.05, dims = 1:12)

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'RNA_snn_res.0.7', repel = TRUE)



```

Add different correlation assignments from the 0.553 threshold and visualize on UMAP from the low threshold for CAM

```{r, fig.width=5}

seu <- AddMetaData(object=seu, metadata= cor1$cell.label, col.name = 'cor.labels55')

# see the labels added
#unique(seu$cor.labels01)

#plot the cluster predictions
#plot_lab_clust(seu, seu$RNA_snn_res.0.7, seu$cor.labels01)

DimPlot(seu, group.by = 'cor.labels55', label = TRUE) + theme(legend.position = "none")


```

Visualize CAM with the signifance 0.553 threshold

```{r, fig.width=5}

seu <- AddMetaData(object=seu, metadata= cor1$cell.label, col.name = 'cor.labels05')

# see the labels added
unique(seu$cor.labels05)


DimPlot(seu, group.by = 'cor.labels05', label = TRUE) + theme(legend.position = "none")


```

Visualize CAM assignments with threshold R = 0.35

```{r, fig.width=5}

seu <- AddMetaData(object=seu, metadata= cor3$cell.label, col.name = 'cor.labels035')

DimPlot(seu, group.by = 'cor.labels035', label = TRUE) + theme(legend.position = "none")

# removing the combined cell labels that are low frequency will improve visualization
# Check if the cell type has a frequency less than 500 and assign new label accordingly
# Calculate cell frequencies
cell.freq <- table(seu$cor.labels035)

# Filter cell labels with at least 400 cells
selected_labels <- names(cell.freq[cell.freq >= 500])

df.meta <- as.data.frame(seu$cor.labels035)
# Create a new column 'cell.label.ft' with labels over 500 cells, and 'NA' for others
df.meta$cell.label.ft <- df.meta$`seu$cor.labels035`
df.meta$cell.label.ft <- ifelse(df.meta$cell.label.ft %in% selected_labels, df.meta$cell.label.ft, NA)

seu <- AddMetaData(object=seu, metadata= df.meta$cell.label.ft, col.name = 'cor.labels035ft')
unique(seu$cor.labels035ft)

DimPlot(seu, group.by = 'cor.labels035ft', label = TRUE, 
        label.size = 8, repel = TRUE) + theme(legend.position = "none")


```

Get the top assigned cells per cluster for each threshold

```{r}
cor.ann.035 <- get_annotation(seu, seu.cluster = seu$RNA_snn_res.0.7, 
                          seu.label = seu$cor.labels035, top_n = 3, 
                          filter_out = c("Unknown","unknown","Mixed", 
                                         "unassigned","Unassigned"), 
                          Label = "CAM")
cor.ann.01 <- get_annotation(seu, seu.cluster = seu$RNA_snn_res.0.7, 
                          seu.label = seu$cor.labels01, top_n = 3, 
                          filter_out = c("Unknown","unknown","Mixed","
                                         unassigned","Unassigned"), 
                          Label = "CAM")
cor.ann.05 <- get_annotation(seu, seu.cluster = seu$RNA_snn_res.0.7, 
                          seu.label = seu$cor.labels05, top_n = 3, 
                          filter_out = c("Unknown","unknown","Mixed",
                                         "unassigned","Unassigned"), 
                          Label = "CAM")


```

Visualize Marker expression

```{r}
length(unique(seu$RNA_snn_res.0.7))
# 19
# if we want to plot by cluster we need a vector of from 0 to the n-1 clusters
cluster.num <- c(0:18)

plotmean(plot_type = 'heatmap',seu = seu, group = 'RNA_snn_res.0.7', markers = AB, 
                     var_names = cluster.num, slot = 'scale.data', xlab = "Cluster",
         ylab = "Antibody")


```
Feature plots

```{r}
Idents(seu) <- "RNA_snn_res.0.7"

for (i in AB) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}


```


Annotate the cells

```{r}

# annotate cells 
Idents(seu) <- "RNA_snn_res.0.7"
cluster.ids <- c("Unassigned","Glial-lineage","Neurons 1","Radial Glia 1",
            "Radial Glia 3","Epithelial","Neurons 2",
            "Astrocytes 1","Astrocytes 1",
                 "Astrocytes 2","Neurons 2","NPC","Radial Glia 1",
            "Radial Glia 2",
                 "Endothelial","Oligodendrocytes","OPC","Stem cell like",
                 "NPC")

seu <- Celltype::annotate(seu, annotations = cluster.ids, to_label = "RNA_snn_res.0.7",
                annotation_name = "labels"
                )

Idents(seu) <- "labels"
# there are 16 levels in the cell type annotations with major groups

# change the order of the cell types on the legend of the umap
cell.type.order <- c("Astrocytes 1", "Astrocytes 2","Radial Glia 1","Radial Glia 2",
                     "Radial Glia 3","Glial-lineage",
                     "Epithelial","Endothelial",
                     "Neurons 1","Neurons 2","NPC",
                     "Oligodendrocytes","OPC","Stem cell like","Unassigned")
#cell.type.order <- rev(cell.type.order)

# colour order to match cell type order
clust.colours <- c("chocolate2","darkorange","pink","plum1","lightpink3",
                   "mistyrose2",
                   "steelblue3","deepskyblue",
                   "mediumpurple1","purple","plum3",
                   "seagreen3","olivedrab4","tomato3","burlywood3")
                   

# Figure 3D UMAP with annotated clusters
# use PDF for figure for correct resolution
DimPlot(seu, order = rev(cell.type.order), cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = FALSE) 

pdf("UMAPlabelled9000.noraster.pdf",width = 9, height = 5)
 DimPlot(seu, order = rev(cell.type.order), cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = FALSE) +
   theme(legend.text = element_text(size=16), axis.title.y = element_text(size=16), 
         axis.title.x = element_text(size=16), axis.text.y = element_text(size =16),
        axis.text.x = element_text(size =16))
dev.off()



######## figure 3E heatmap of the 

# reorder the bars to match the UMAP
levels(seu) <- c("Astrocytes 1", "Astrocytes 2","Radial Glia 1","Radial Glia 2",
                     "Radial Glia 3","Glial-lineage",
                     "Epithelial","Endothelial",
                     "Neurons 1","Neurons 2","NPC",
                     "Oligodendrocytes","OPC","Stem cell like","Unassigned")

pdf("HM9000_fig3E.pdf",width = 8.5, height = 5)
DoHeatmap(seu, features = AB, size= 6,slot = "scale.data", group.colors = clust.colours, disp.max = 2, disp.min = -1.5,
          angle = 90) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 15))
dev.off()

```

Train a Random Forest classifier

```{r}

# get the names of hte meta data to know the annotations to call
colnames(seu@meta.data)

DimPlot(seu, group.by = "labels")

```



Train a Random Forest classifier 

Using too few features (mytry), too low of starting nodes (tree complexity), too few nodes (max nodes) and too few trees can all lead to over fitting.  The num_folds variable provides a cross validation within the parameter search to help avoid over fitting. 

```{r}

# 
markers <- rev(c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4"))
# mytry is for the number of features to be tested as input
# here we have a max of 13 putting all the features doesn't work to create a RFM because all the trees will be the same.

# To force an assignment I am removing the "Unassigned" cells in this model.

Idents(seu) <- "labels"
seu.sub <- subset(seu, idents = "Unassigned", invert = TRUE)


#Remove the "Unassigned" level from the Idents metadata
Idents(seu.sub) <- Idents(seu.sub)[Idents(seu.sub) != "Unassigned"]
#Update the labels in the subset object
seu.sub$labels <- Idents(seu.sub)


seu.sub <- subset(seu, downsample = 1000)
# we can down sample in the function but this will be random across all cell types.  Downsampling based by cell type will down sample to 1000 of each cell type or the total if there is less than 1000.  This will make the cell types more balanced

rf <- RFM_train(seurate_object = seu.sub,
                      markers, 
                      annotations = seu.sub$labels,
                      num_folds = 3,
                      downsample = 'none',
                      seed = 222,
                      mytry = c(6:7),
                      maxnodes = c(15:18),
                      trees = c(1800,2000),
                      cores_to_use = 4)



```

Save the model
 
```{r}
saveRDS(rf,"/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/NatMethodJuneSubmission/RFM/MBOsub_trainedRFMnewlabelsAu17.Rds")
```
 
 