---
title: "Figure 2"
output: github_document
---


This workbook contains:

1. Pre-processing of the individual 2D cultures
2. Create of figures in figure 2.


#Read in the 2D data

Load required libraries
```{r}
# set up environment / load libraries
library(Seurat)
library(dplyr)
library(ggplot2)
library(CelltypeR)

```

Read in the live gated single cell fsc files

```{r}
input_path <- "/Users/rhalenathomas/GITHUB/CelltypeR/FlowCytometry_Data/2Dcultures"

flowset <- fsc_to_fs(input_path, downsample = "none")
# down sample can be a number, 'none' or 'min'
# look at file names and rename with shorter sample names
library(flowWorkspace)  # this library has the function "sampleNames"
sampleNames(flowset)



```

Rename frames in the list
```{r}
sampleNames(flowset) <- sampleNames(flowset) <- c("Astrocytes1","Neurons","NPC1","Oligodendrocyte-m","Astrocyte2","IPSC","NPC2","Oligodendrocyte-e")
sampleNames(flowset)
```


```{r}
plotdensity_flowset(flowset)
flowset_biexp <- harmonize(flowset, processing = 'biexp')

# we need to view the biex transform to see which measurements have two peaks
plotdensity_flowset(flowset_biexp)
```
Align and retro transform, make a data frame and save a csv

```{r}
flowset_retro <- harmonize(flowset, processing = 'retro', 
                           two_peaks = c(7,9:19),
                       one_peak = c(1:6,8), threshold = 0.01)


# now create the data frame from the flowset object using the CelltypeR function flowset_to_csv
df <- flowset_to_csv(flowset_retro)

write.csv(df, "2Dcells_surfaceflowset.csv")

```

# Create the Seurat object and cluster

```{r}
markers <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")
seu <- make_seu(df, AB_vector = markers)
Idents(seu) <- "Sample"
levels(seu)
celltypes <- c("Astrocytes","Neurons","NPC","Oligodendrocytes","Astrocytes","IPSC","NPC","Oligodendrocytes")
seu <- CelltypeR::annotate(seu1, annotations = celltypes, to_label = 'Sample', annotation_name = 'Batch')
Idents(seu) <- 'Batch'
seu <- subset(seu, downsample = 10000)
# note that Batch here refers to the culture types

seu <- get_clusters(seu, method = "louvain", k = 25,
                    resolution = 0.25, plots = FALSE)
DimPlot(seu, group.by = 'RNA_snn_res.0.25', repel = TRUE, raster = FALSE, pt.size = 0.25, label = TRUE) 



```

Save Seurat object for later
```{r}
saveRDS(seu, "Seurat2DcellCultures.RDS")
#seu <- readRDS("Seurat2DcellCultures.RDS")

```

Make heatmap of expression by cell type sample

```{r}

# this is the seurat plotting function with bar colours and scale colours assigned.
hm <- DoHeatmap(seu, group.by = "Batch", features = markers, size= 4,slot = "scale.data", group.colors = c("dodgerblue2","bisque","seagreen3","red3","mediumpurple3"), disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 12))
hm

pdf("Heatmap_2DcellSamples.pdf",width =5.5, height = 4)
hm
dev.off()



```


UMAPS of 2D cell types

```{r}
# plot the UMAP

cell.order <- rev(c("Astrocytes","IPSC","Neurons","NPC","Oligo"))
Idents(seu) <- "Batch"
DimPlot(seu, repel = TRUE, label = TRUE, cols = c("dodgerblue2","bisque","seagreen3","red3","mediumpurple3"), order = cell.order, shuffle = TRUE ) 



pdf("kn25UMAPcelltypes2D.pdf",width =6, height = 4)
DimPlot(seu,label = FALSE, cols = c("dodgerblue2","bisque","seagreen3","red3","mediumpurple3"),
        order = cell.order, shuffle = TRUE) 
dev.off()

```

```{r, fig.width=8}
DimPlot(seu, repel = TRUE, label = TRUE, order = cell.order, shuffle = TRUE, split.by = "Batch", group.by = "RNA_snn_res.0.25") 

Idents(seu) <- "RNA_snn_res.0.25"
pdf("kn25UMAPcelltypes2DSplitBy.pdf",width =20, height = 4)
DimPlot(seu, repel = TRUE, label = TRUE, order = cell.order, shuffle = TRUE, split.by = "Batch") 
dev.off()
Idents(seu) <- "Batch"
pdf("kn25UMAPcelltypes2DSplitByCellColours.pdf",width =20, height = 4)
DimPlot(seu, repel = TRUE, label = TRUE, order = cell.order, shuffle = TRUE, split.by = "Batch",cols = c("royalblue3","bisque3","springgreen4","red3","mediumpurple3")) 
dev.off()


```


Visualize the clusters

```{r}
markers <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

for (i in markers) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}

# save figure
pdf("kn25UMAPmarkers.pdf",width =20, height = 16)

FeaturePlot(seu, features =markers, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE, raster = FALSE, ncol = 4)
dev.off()

```


Correlation predictions

```{r}

reference_path <- "GITHUB/CelltypeR/ExampleOuts/RefrecneMatrices/FinalReferenceMatrix.csv"
reference_data <- read.csv(reference_path)

rownames(reference_data) <- reference_data$X
reference_data <- reference_data %>% select(-"X")
#
head(reference_data)
colnames(reference_data) <- c("Astrocytes", "Endothelial", "Epithelial", "Neurons","NPC", "OPC",  "Oligodendrocyte", "Radial Glia", "Stem Like")

df_ref <- as.data.frame(t(reference_data))
df_ref$X <- rownames(df_ref)

# rows are cell types in the reference matrix and Markers are the columns
# in the input markers are the column.  


input_df <- read.csv("2DcellsSelectflowset.csv")
dim(input_df)
colnames(input_df)
class(input_df)

# the reference matrix and the test matrix need to both have the Markes as column
# the ref matrix has a column of Cell types "X"
# the min_corr cut-off will effect the amount of assigned cells

cor1 <- find_correlation(test = input_df, reference = df_ref,
                        min_corr = 0.35, min_diff = 0.05)

# add the correlation prediction for each cell
# thresh 0.35
seu <- AddMetaData(object=seu, metadata=cor1$cell.label, col.name = 'cor.labels1')


```


```{r}
#table(seu$cor.labels1)
table(seu$cor.labels4)

freq.cor4 <- as.data.frame(table(seu$cor.labels4))

```


Have a look at the correlations

```{r, fig.width=10}
Idents(seu) <- 'cor.labels'
DimPlot(seu) # see all correlation predictions 
# top correlations for each cluster
cor.ann.035 <- get_annotation(seu, seu.cluster = seu$RNA_snn_res.0.25, 
                          seu.label = seu$cor.labels, top_n = 3, 
                          filter_out = c("Unknown","unknown","Mixed", 
                                         "unassigned","Unassigned"), 
                          Label = "CAM")

# top 5 per groups
t.lables <- as.data.frame(table(seu$RNA_snn_res.0.25, seu$cor.labels))
  t.lables$Freq <- as.double(t.lables$Freq)
  colnames(t.lables) <- c("Cluster", "Label","Freq")
  top.labs <- t.lables  %>% group_by(Cluster) %>% top_n(5, Freq)
  sort.tops <- top.labs %>% as.data.frame() %>% arrange(desc(Freq))  %>% arrange(Cluster)
  print(sort.tops)
  cor.ann.035

  # make a bar chart of the cell type per cluster
  plot_lab_clust(seu, seu$RNA_snn_res.0.25, seu$cor.labels)

  
  pdf("BarchartCorbyClusterThresh035.pdf", width = 8, height = 4)
  plot_lab_clust(seu, seu$RNA_snn_res.0.25, seu$cor.labels)
  DimPlot(seu)
  dev.off()

```


```{r}

seu <- CelltypeR::annotate(seu, annotations = cor.ann.035$CAM, to_label = "RNA_snn_res.0.25",
                annotation_name = "CAM035")
DimPlot(seu, group.by = "CAM035", label = TRUE)

```


Label clusters and plot UMAP again

```{r}


Idents(seu) <- "RNA_snn_res.0.25"

cluster.ids2 <- c("Astrocytes","Mix","Stem Cells",
                 "NPC","Oligo",
                 "NPC-Radial Glia",
                 "Neurons","Radial Glia",
                 "Astrocytes-Stem","Epithelial-Endothelial",
                 "Endothelial")

cellorder2 <- c("Astrocytes","Astrocytes-Stem",
               "Epithelial-Endothelial","Endothelial","Mix",
               "Neurons",
                "NPC","NPC-Radial Glia","Oligo",
               "Radial Glia", "Stem Cells")


seu <- annotate(seu, annotations = cluster.ids2, 
  to_label = 'RNA_snn_res.0.25',
  annotation_name = "Subgroups2")
# plot the UMAP 
Idents(seu) <- 'Subgroups2'
length(levels(seu))


#UMAP
Idents(seu) <- 'Subgroups2'
DimPlot(seu, repel = TRUE, label = TRUE,
        cols = c("dodgerblue2","royalblue3",
                 "cadetblue3","darkturquoise","wheat4",
                 "seagreen3","red3","indianred1",
                 "mediumpurple3","pink",
                 "bisque"), order = rev(cellorder2), 
        shuffle = TRUE) 



pdf("kn25UMAPSubgroups2Dcells.pdf",width =6, height = 4)
DimPlot(seu, repel = TRUE, label = FALSE,
        cols = c("dodgerblue2","royalblue3",
                 "cadetblue3","darkturquoise","wheat4",
                 "seagreen3","red3","indianred1",
                 "mediumpurple3","pink",
                 "bisque"),
                order = rev(cellorder2), shuffle = TRUE) 
dev.off()



```


Mean expression per cluster

```{r}
library(data.table)



cellorder2 <- c("Astrocytes","Astrocytes-Stem",
               "Epithelial-Endothelial","Endothelial","Mix",
               "Neurons",
                "NPC","NPC-Radial Glia","Oligo",
               "Radial Glia", "Stem Cells")


cluster.ids2 <- c("Astrocytes","Mix","Stem Cells",
                 "NPC","Oligo",
                 "NPC-Radial Glia",
                 "Neurons","Radial Glia",
                 "Astrocytes-Stem","Epithelial-Endothelial",
                 "Endothelial")


markers <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")
hm <- plotmean(plot_type = 'heatmap',seu, group = "Subgroups2", markers = markers, 
         var_names = cluster.ids2, slot = 'scale.data',
                     xlab = 'Cell Types', ylab = 'Markers', 
         cluster_order = cellorder2, 
         marker_order = rev(markers), high_colour = "darkblue",
         mid_colour = "slateblue", low_colour = "ivory")+
  theme(panel.border = element_blank(), 
           panel.grid = element_blank())
hm


#pdf("heatmap_2DexpbySubgroup2.pdf",width = 4.5, height = 4)
#print(hm)
#dev.off()

hm




```




