---
title: "Figure 7"
output: github_document
---

1. Data preprocessing: put all timepoints and replicates into one folder
a. Read FlowJo files into R.
b. Create a data frame with intensity measurements for each marker for all          samples within the experiment to be analyzed.  
- downsample taking a 2000 cells from each sample to speed up computation.
c. Harmonize data 
d. Create dataframe
e. Create a Seurat object

2. Cluster data
3. Annotate clusters with cell types.
a. Predict cell types by correlation matrix
b. Visualize expression by cluster for manual annotation
c. Add annotations to Seurat object
c. Train RandomForest classifier.

4. Reprocess with all cell in samples
a. preprocess
b. cluster
c. annotate (now seurat label transfer and random forest can be used.)

5. Compare cell type proportions and expression levels across the time course. 

```{r}
# load necessary libraries 
library(Seurat)
library(dplyr) 
library(ggplot2)
library(CelltypeR)

```


# Preprocessing

Read in the flow data
This data should be the gated live cells.  
All samples need to be in one folder.
Here we have 

```{r}

input_path <- "/Users/rhalenathomas/GITHUB/CelltypeR/FlowCytometry_Data/AIW_timeCourse"

# 1.a Read in FlowJo Files 

# down sample to take cells from each of 4 time points and 4 replicates
flowset <- fsc_to_fs(input_path, downsample = 2000)
# down sample can be a number, 'none' or 'min'

# look at file names and rename with shorter sample names
sampleNames(flowset) # function in flowCore

```

Rename samples with shorter names

```{r}
sampleNames(flowset) <- sampleNames(flowset) <- c("AIW_150_R1","AIW_150_R2","AIW_150_R3","AIW_150_R4",
  "AIW_30_R1","AIW_30_R2","AIW_30_R3","AIW_30_R4",
  "AIW_60_R1","AIW_60_R2","AIW_60_R3","AIW_60_R4",
  "AIW_100_R1","AIW_100_R2","AIW_100_R3","AIW_100_R4"
                                                  )
sampleNames(flowset)
```

Harmonize data to remove batch or technical variation

This requires us to look and see where there are two peaks to align. We need to visualize the peaks of the transformed data before aligning.

```{r}

# we can decided what level of processing to choose with the argument 'processing'
# biexp only applies a biexponential transformation
# align applies biexp transform and then aligns peaks
# retro (default), transforms, aligns and then reverse transforms
flowset_biexp <- harmonize(flowset, processing = 'biexp')
# we can visualize the peaks to see where there are two peaks for alignment

# we need to enter the column index for which peaks to align, the alignment for one or two peaks is not the same. 
#plotdensity_flowset(flowset)

p1 <- plotdensity_flowset(flowset_biexp) # to see the peaks
p1
# if there are two peaks assign as two peaks and one peak as one in the code chunk below

flowset_align <- harmonize(flowset, processing = 'align', 
                           two_peaks = c(9:11,15,17:21,25:42),
                       one_peak = c(1:8,12:14,16,24,43), threshold = 0.01)

flowset_retro <- harmonize(flowset, processing = 'retro', 
                           two_peaks = c(9:11,15,17:21,25:42),
                       one_peak = c(1:8,12:14,16,24,43), threshold = 0.01)

df <- flowset_to_csv(flowset_retro)



```
Create the Seurat object 

```{r}

# antibody features in order to appear on the plots
AB <- c("TH","CD24","CD56","CD29","CD15","CD184","CD133","SSEA4","CD44","CD49f","CD140a")

df_select <- df %>% select(c("TH","CD24","CD56","CD29","CD15","CD184","CD133","SSEA4","CD44","CD49f","CD140a","Sample"))

seu <- make_seu(df_select, AB_vector = AB)

```


# Cluster
Explore conditions

```{r}
# note the dimensions (number of PC to include needs to be one less than the number in the antibody panel)
explore_param(input = seu, 
                          cluster_method = "louvain", 
                          df_input = df.input, 
                          flow_k = NULL, 
                          pheno_lou_kn = c(80,120,200), 
                          lou_resolution = c(0.6), 
                          pcdim = 1:10,
                          run.plot = TRUE, 
                          run.stats = FALSE, 
                          save_to = NULL)
# this function does not add the clustering to the seurat object it is to explore and run stats.  If you choose save = output_path then you will save a seurat object with the clustering, one for each kn.


```


After checking parameters run cluster function to make add the clustering information into the 

```{r}

seu <- get_clusters(seu, method = "louvain",
                         df_input = df.input,
                         k = 80,
                         resolution = c(1),
                         plots = TRUE,
                         save_plots = FALSE)


```


# Annotate clusters
1. Visualization for manual annotation. - output by clustering function
2. CAM (Correlation assignment model) - requires reference matrix
3. RFM (Random Forest Model) - requires annotated matching flow dataset
4. Seurat label transfer - requires annotated matching flow data in a seurat object

Visualize expression on UMAP and with heat maps

```{r}

AB <- c("TH","CD24","CD56","CD29","CD15","CD184","CD133","SSEA4","CD44","CD49f","CD140a")
# the cluster labels will match the active ident
Idents(seu) <- "RNA_snn_res.1"
# this will let us see one at at time
for (i in AB) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q97', label = TRUE))
}


```

Visualize your reference matrix

```{r}
df <- read.csv("GITHUB/CelltypeR/ExampleOuts/ReferenceMatrices/ReferenceMatrixPanel2.csv")

# we need to remove the marker column and add these are row names
rownames(df) <- df$X
df <- df %>% select(-"X")
# heatmap function take a matix
mat <- as.matrix(df)

heatmap(mat, 
        Rowv = NA,  # Don't cluster rows
        Colv = NA,  # Don't cluster columns
        col = colorRampPalette(c("white", "blue"))(100),  # Define a color palette
        main = "Reference Matrix")



```

```{r, fig.width=3}
colnames(seu@meta.data)
DimPlot(seu, group.by = "RNA_snn_res.1", label=TRUE, label.size = 6)
DimPlot(seu, group.by = "Days_in_FD", label=TRUE, label.size = 6)
DimPlot(seu, group.by = "Replicate", label=TRUE, label.size = 6)


```


Visualization of expression values by heatmap

```{r}

# summary heat map
# use function plotmean
# we need to know how many clusters we have.  They will be index 0 to n-1
# you need to pick the length to annotate

length(unique(seu$RNA_snn_res.1))
# 15
# if we want to plot by cluster we need a vector of from 0 to the n-1 clusters
cluster.num <- c(0:17)

plotmean(plot_type = 'heatmap',seu = seu, group = 'RNA_snn_res.1', markers = AB, 
                     var_names = cluster.num, slot = 'scale.data', xlab = "Cluster",
         ylab = "Antibody")


```

```{r, fig.width=3}

DimPlot(seu, group.by = "Sample")
```


Add annotations based on expression

```{r}
# predict cell type from looking at expression patterns
myann <- c("precursors","neuralblast","neuron1","astrocyte1","stem-like",
           "epithelial","astrocyte-early","astrocyte2","DAneuron","RG-astro",
           "DAneuron2",
           "Neuron_early","RG","Glia","NPC","OPC","neuralstem","stem-like2")

man.ann <- data.frame(Cluster = c(0:17), MyAnn = myann)
man.ann

man.ann$Cluster <- as.factor(man.ann$Cluster)
man.ann$MyAnn <- as.factor(man.ann$MyAnn)
man.ann

```

Predict cell annotations with CAM (Correlations assignment method)

```{r}

reference_path2 <- "/GITHUB/CelltypeR/ExampleOuts/ReferenceMatrices/ReferenceMatrixPanel2Timecourse.csv"
reference_data2 <- read.csv(reference_path2)
# the reference matrix need to be in the format cell types as rows and markers as columns
# there is a column X with the cell type names
df1 <- reference_data2
rownames(df1) <- df1$X # add row names (these are the markers)
df1 <- df1 %>% select(-"X") # remove the column with the marker names

cor <- find_correlation(test = input_df, 
                             reference = df1, 
                             min_corr = 0.35, 
                             min_diff = 0.05)

# creates a dataframe with cor1 cor2 and predicted cell type label

cor2 <- find_correlation(test = input_df, 
                             reference = df1, 
                             min_corr = 0.55, 
                             min_diff = 0.05)


```

Visualize the CAM results

```{r}
#dim(cor)
#colnames(cor)

plot_corr(cor, threshold = 0.35, min_cells = 400)

```

```{r}
plot_corr(cor2, threshold = 0.55, min_cells = 400)
```


```{r, fig.width= 4}

# add the correlation predictions to the meta data
seu <- AddMetaData(object=seu, metadata=cor$cell.label, col.name = 'cor.labels3505')
# see the labels added
#unique(seu$cor.labels3505)

# add the correlation predictions to the meta data
seu <- AddMetaData(object=seu, metadata=cor2$cell.label, col.name = 'cor.labels5505')
# see the labels added
#unique(seu$cor.labels5505)

# plot the cluster predictions
plot_lab_clust(seu, seu$RNA_snn_res.1, seu$cor.labels5505)

```

Run get annotations function to return a vector of annotation in the order of the clusters.

```{r}

cor.ann <- get_annotation(seu, seu.cluster = seu$RNA_snn_res.1, 
                          seu.label = seu$cor.labels3505, top_n = 3, 
                          filter_out = c("Unknown","unknown","Mixed","unassigned","Unassigned"), 
                          Label = "CAM3505")
cor.ann
dim(cor.ann)

unique(cor.ann$CAM)
length(cor.ann$CAM)

```

```{r, fig.width= 3}

ann <- c("NPCDA1", "Astrocyte1","Neuron1", "Astrocyte2",
         "RG1","RG2", "Astrocyte3", "Astrocyte4",
         "NeuronDA","Astrocyte5","NPCDA2","Neuron2",
         "RadialGliaDiv","RG3","Neuron4","opc", "RG4", "StemCell")


seu <- annotate(seu,annotations = cor.ann$CAM3505, 
                to_label = "RNA_snn_res.1", 
                annotation_name = "CAM35")


DimPlot(seu, group.by = "CAM35", label = TRUE, label.size = 6)


```


```{r, fig.width=3}
cor.ann <- get_annotation(seu.t, seu.cluster = seu.t$RNA_snn_res.1, 
                          seu.label = seu.t$cor.labels5505, top_n = 3, 
                          filter_out = c("Unknown","unknown","Mixed","unassigned","Unassigned"), 
                          Label = "CAM5505")
cor.ann
dim(cor.ann)

unique(cor.ann$CAM5505)
length(cor.ann$CAM5505)

seu <- annotate(seu,annotations = cor.ann$CAM5505, 
                to_label = "RNA_snn_res.1", 
                annotation_name = "CAM5505")


DimPlot(seu, group.by = "CAM5505", label = TRUE)


```



We do not have a trained Random Forest model or a labelled Seurat object to compare for this antibody set. We can test RFM prediction and Seurat transfer using the older organoids with the original panel.However only the overlapping antibodies can be used. 

```{r}
# which antibodies overlap between the two panels

panel1 <-  c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

panel2 <- c("TH","CD24","CD56","CD29","CD15","CD184","CD133","SSEA4","CD44","CD49f","CD140a")

overlap_Markers <- intersect(panel1, panel2)
print(overlap_Markers)
  
  
```

Random Forest Model predictions with only overlapping markers.

```{r}

# you must have a saved trained model from a data object annotated from the same markers
# I have trained a model with only the overlapping antibodies

overlabling_markers <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD44", "CD140a")

# Random Forest Model was trained on HPC.  However with so few features the predictive ability will be limited. 

rf <- readRDS("/RFM_all9hMOsOverlapAB.RDS")


rfm.pred <- RFM_predict(seu.t, rf)


seu.t <- AddMetaData(object=seu.t, metadata=rfm.pred$Prediction, col.name = 'rfm.labels')

# check that the data is added 
table(seu.t$rfm.labels)



```

Get the annotation by cluster for the RFM

```{r}

rfm.ann <- get_annotation(seu.t, seu.t$RNA_snn_res.1,seu.t$rfm.labels, 
               top_n = 3, filter_out = c("unknown","Unknown","Mixed","Mix", "unassigned"), Label = "RFM")
```

Plot RFM predictions

```{r}

plot_lab_clust(seu.t, seu.cluster = seu.t$RNA_snn_res.1, seu.labels = seu.t$rfm.labels, filter_out = c("unknown","Unknown","Mixed"))


```

Predicting cell types with Seurat label transfer using anchors


```{r}

# takes in a seurat object with the labels added 
# makes a dataframe with the count of predicted labels for each cluster
# input seurat object with the predicted labels in the meta data
# input the clusters meta data slot to be labels

#need reference data object with labels
seu.r<- readRDS("All9MOannaoteAug.RDS") 

Idents(seu.r) <- 'Celltypes'
seu.r <- subset(seu.r, downsample = 5000)

overlabling_markers <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD44", "CD140a")
seu.t <- seurat_predict(seu.t, seu.r, refdata = seu.r$Celltypes,
                      ref_id = 'Celltypes', down.sample = "none", markers = overlabling_markers)

DimPlot(seu.t, group.by = "seu.pred", reduction = "umap")

```



```{r}

# plot the seurat anchor predictions
# get the annotation table for the seurat anchor predictions 

plot_lab_clust(seu.t, seu.t$RNA_snn_res.1, seu.t$seu.pred)

# to not filter anything use c()
seu.ann <- get_annotation(seu.t, seu.t$RNA_snn_res.1, seu.t$seu.pred, 
               top_n = 3, filter_out = c("Unknown","Mixed","unassigned",
                                         "Unassigned"), Label = "Seurat")



```

```{r, fig.width=3}
# add the annotations from the seurat predictions

seu.t <- annotate(seu.t, annotations = seu.ann$Seurat, 
                to_label = "RNA_snn_res.1", 
                annotation_name = "Seu")


DimPlot(seu.t, group.by = "Seu", label = TRUE, label.size = 6)

  
```


Make a table with the top predicted cell by cluster

```{r}
an.list <- list(cor.ann, rfm.ann, seu.ann)
df.an <- annotate_df(an.list)
df.an
write.csv(df.an, "annotationPredsTop1.csv")

```

Get a consensus of cluster annotations, Add the annotations to the seurat object


```{r}

# make a list of data frames
# all the dataframes need to be as.factor
ann.list <- list(cor.ann,rfm.ann,seu.ann.ft, man.ann)

# annotate the seurat object

seu <- cluster_annotate(seu, ann.list = ann.list, 
                        annotation_name ="CellType", 
                        to_label = "RNA_snn_res.1")

DimPlot(seu, group.by = "CellType")

# unfortunately for the consensus to work we need the cell type names to be identical

```

Add annotations

```{r}

seu <- annotate(seu, annotations = new_annotations, to_label = "RNA_snn_res.1", annotation_name = "CellTypes")

DimPlot(seu, group.by = "CellTypes", label = TRUE, repel = TRUE)

new_annotations <- c("earlyDA-NPC","RG1a-astro","Neurons1","Astrocytes1-RG1a",
                     "stem-like-RG2","RG3","RG1","Astrocytes2",
                     "DAneurons1","RG-astro","earlyDA-NPC2","NPC",
                     "RGdiv-NPC-RG1a","RG1a-astro1", "Neurons1b","OPC-like",
                     "precursors","Neurons2")

seu.t <- annotate(seu.t, annotations = new_annotations, to_label = "RNA_snn_res.1", annotation_name = "CelltypesNew")

DimPlot(seu.t, group.by = "CelltypesNew", label = TRUE)

new_annotations <- c("DA-NPC","Astrocytes 1","Neurons 1","RG1a-Astrocytes 1",
                     "RadialGlia 2","RadialGlia 3","RadialGlia 1","Astrocytes 2",
                     "DA-neurons 1","RG1a-Astrocytes 1","DA-NPC","Neurons 1",
                     "Stem-cell-like","RadialGlia 1a", "NPC","OPC-like",
                     "RG-oligo","Neurons 2")

seu.t <- annotate(seu.t, annotations = new_annotations, to_label = "RNA_snn_res.1", annotation_name = "CelltypesMainNew")

DimPlot(seu.t, group.by = "CelltypesMainNew", label = TRUE)

```


```{r}
colnames(seu.t@meta.data)
```

# Compare groups

We first need to add the variables into the seurat object that we want to compare.

```{r}
# see the sample names and order to add the meta data

# set the ident to sample names
Idents(seu.t) <- 'Sample'
levels(seu.t)

Age <- c("155","155","155","150",
         "30","30","30","30",
         "60","60","60","60",
         "100","100","100","100")
Replicate <- c("R1","R2","R3","R4",
               "R1","R2","R3","R4",
               "R1","R2","R3","R4",
               "R1","R2","R3","R4")


# vector with the new names - you need this vector from me
cluster.ids <- Age

names(cluster.ids) <- levels(seu)    # get the levels
seu <- RenameIdents(seu, cluster.ids) # rename  
seu$Days_in_FD <- Idents(seu)   # add a new dataslot

# Replicate
Idents(seu) <- "Sample"
cluster.ids <- Replicate
# vector with the new names - you need this vector from me

names(cluster.ids) <- levels(seu)    # get the levels
seu <- RenameIdents(seu, cluster.ids) # rename  
seu$Replicate <- Idents(seu)   # add a new dataslot

# change the levels in the time course so that are in the order of actual time for plotting

# rename with exact time points

Idents(seu.t) <- "Days_in_FD"
levels(seu.t)
days <- c("38","63","98","155")


seu.t <- annotate(seu.t, annotations = days, to_label = "Days_in_FD", 
                annotation_name = "Days_in_FD")

seu.t$Days_in_FD <- factor(seu.t$Days_in_FD, levels = c("38","63","98","155"))

Idents(seu.t) <- "Days_in_FD"
DimPlot(seu.t)

```

See the data split by days in FD

```{r}
DimPlot(seu.t, split.by = "Days_in_FD", ncol = 2, label = TRUE)
DimPlot(seu.t, split.by = "Replicate", ncol = 2, label = TRUE, label.size = 2)
```

```{r}
# set up colours to match Fig 4,5 where possible


cell.order <- c("Astrocytes 1","Astrocytes 2","RG1a-Astrocytes 1","RadialGlia 1",
                "RadialGlia 1a","RadialGlia 2","RadialGlia 3",
                "DA-neurons 1","Neurons 1","Neurons 2","DA-NPC","NPC",
                "OPC-like",
                "RG-oligo",
                "Stem-cell-like")

# add TH specific
clust.colours <- c("chocolate2",# Astrocytes 1
                   "darkorange", # Astrocytes 2
                   "salmon",  # RG1a-Astrocytes
                   "pink", # RG1
                   "deeppink",# RG1a
                   "plum1", #RG2
                   "lightpink3",# RG3
                   "slateblue", #"DA-Neuron"
                   "mediumpurple1",# neurons 1
                   "purple",# Neurons2
                   #"mistyrose2", #glialineage
                   #"steelblue3",# epithelial
                   #"deepskyblue", # endothelial
                    "thistle2",   #DA-NPC
                   "plum3", # NPC
                   #"mediumslateblue", # Neural lineage
                   # "seagreen3",#Oligo
                   # "olivedrab4", # OPC
                   "darkseagreen3",#OPC like
                   "yellowgreen", # RG-oligo
                   "tomato3",# stem like - NPC-RG-div
                   "burlywood3" #extra 
                   )

```

UMAP with colours

```{r}

DimPlot(seu.t, cols = clust.colours, shuffle = TRUE, 
        raster=FALSE, pt.size = 0.1, label = TRUE, 
        group.by = "CelltypesMainNew",
        order = rev(cell.order)) +
theme(legend.text = element_text(size=16), axis.title.y = element_text(size=16), 
         axis.title.x = element_text(size=16), axis.text.y = element_text(size =16),
        axis.text.x = element_text(size =16))

# save the plot
pdf("UMAPtimeCourse.pdf",width = 9, height = 4.5)
DimPlot(seu.t, cols = clust.colours, shuffle = TRUE, 
        raster=FALSE, pt.size = 0.1, label = FALSE, 
        group.by = "CelltypesMainNew",
        order = rev(cell.order)) +
theme(legend.text = element_text(size=16), axis.title.y = element_text(size=16), 
         axis.title.x = element_text(size=16), axis.text.y = element_text(size =16),
        axis.text.x = element_text(size =16))
dev.off()


```

Plots some variables to look for differences between groups

```{r}

# add colours from above
# add new cell order 

seu.t$CelltypesMainNew <- factor(seu.t$CelltypesMainNew, levels = cell.order)


# same thing with custom colours
proportionplots(seu.t,seu.var = seu.t$Days_in_FD, seu.lable = seu.t$CelltypesMainNew, groups = "Days_in_FD", my_colours = clust.colours)

# save the plot
pdf("BarcharttimeCourse.pdf",width = 5, height = 4)
proportionplots(seu.t,seu.var = seu.t$Days_in_FD, seu.lable = seu.t$CelltypesMainNew, groups = "Days_in_FD", my_colours = clust.colours)
dev.off()


```

Look at cell type proportions across time and replicates

```{r}
# order the samples
unique(seu.t$Sample)
neworder <- c("AIW_30_R1","AIW_30_R2","AIW_30_R3","AIW_30_R4",
              "AIW_60_R1","AIW_60_R2", "AIW_60_R3","AIW_60_R4",
              "AIW_100_R1","AIW_100_R2","AIW_100_R3", "AIW_100_R4",
              "AIW_150_R1", "AIW_150_R2", "AIW_150_R3", "AIW_150_R4")

seu.t$Sample <- factor(seu.t$Sample, levels = neworder)

proportionplots(seu.t,seu.var = seu.t$Sample, seu.lable = seu.t$CelltypesMainNew, groups = "Days in FD and replicates", my_colours = clust.colours)

#unique(seu.t$sample)
setwd("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/TimeCourseAIW/Analysis/Figure7/")
# save the plot
pdf("BarcharttimeCourseSamples.pdf",width = 6, height = 4)
proportionplots(seu.t,seu.var = seu.t$Sample, seu.lable = seu.t$CelltypesMainNew, groups = "Days in FD and replicates", my_colours = clust.colours)
dev.off()

```

Another visualization over time

```{r}
df <- as.data.frame(proportions(table(seu.t$CelltypesMainNew,seu.t$Days_in_FD,seu.t$Replicate)))
colnames(df) <- c("Celltypes","DaysFD","Replicate","Proportion")

head(df)

df$DaysFD <- as.numeric(as.character(df$DaysFD))

# Create the plot
ggplot(df, aes(x = DaysFD, y = Proportion *1000, color = Celltypes)) +
  geom_point() +
  stat_summary(data = mean_df, aes(x = DaysFD, y = MeanProportion *1000, group = Celltypes), 
               fun = "mean", geom = "line", linetype = "solid", size = 1) +
  labs(title = "Mean Proportion vs. DaysFD",
       x = "DaysFD",
       y = "Percent of total cells",
       color = "Celltypes") +
  scale_color_manual(name = "Celltypes", values = clust.colours) +
  scale_x_discrete() + # Set x-axis limits
  facet_wrap(~Celltypes, scales = "free_y") + # Split by facets
   theme(strip.text.x = element_text(size =8,angle = 0, hjust = 0.5, vjust = 0, margin = margin(b = 6)))

pdf("LinePlotPercentbyTime.pdf", width = 10, height = 7)
ggplot(df, aes(x = DaysFD, y = Proportion *1000, color = Celltypes)) +
  geom_point() +
  stat_summary(data = mean_df, aes(x = DaysFD, y = MeanProportion *1000, group = Celltypes), 
               fun = "mean", geom = "line", linetype = "solid", size = 1) +
  labs(title = "Mean Proportion vs. DaysFD",
       x = "DaysFD",
       y = "Percent of total cells",
       color = "Celltypes") +
  scale_color_manual(name = "Celltypes", values = clust.colours) +
  scale_x_discrete() + # Set x-axis limits
  facet_wrap(~Celltypes, scales = "free_y") + # Split by facets
   theme(strip.text.x = element_text(size =8,angle = 0, hjust = 0.5, vjust = 0, margin = margin(b = 6)))
dev.off()


unique(df$DaysFD)

 
```


```{r}

# anovoa on proportions n=4 replicates
# independent variable is Days in culture with 4 levels
df <- as.data.frame(proportions(table(seu.t$CelltypesMainNew,seu.t$Days_in_FD,seu.t$Replicate)))
colnames(df) <- c("Celltypes","DaysFD","Replicate","Proportion")

two_way <- aov(Proportion ~  Celltypes * DaysFD, data = df)
output <- summary(two_way)
        # now the posthoc test
tukey <- TukeyHSD(two_way)


Anova2 <- as.data.frame(output[[1]])
interaction_df <- as.data.frame(tukey[[3]])
### filter to get the interactions that are interesting

# affect of time 
time_df <- as.data.frame(tukey$DaysFD)
cell_df <- as.data.frame(tukey$Celltypes)
# Assuming df is your dataframe with contrast names in row names
# Create a logical vector indicating whether the cell types match in each row
matches_celltypes <- grepl("^([^:]+):\\d+-\\1:\\d+", rownames(interaction_df))

# Filter the dataframe to keep only rows where cell types match
filtered_df <- as.data.frame(interaction_df[matches_celltypes, ])

# View the filtered dataframe
filtered_df
df <- filtered_df
# add columns for the celltypes and time points
df$Celltype <- sub(":.*", "", rownames(df))
df$time1 <- as.integer(sub(".*:", "", rownames(df)))
df$time2 <- as.integer(sub(".*:", "", sub("-.*", "", rownames(df))))


# Filter the dataframe to keep rows where "p adj" <= 0.05
sig_df <- as.data.frame(df[df$`p adj` <= 0.05, ])

results.list <- list("anova" = Anova2, "Tukey_DaysFD" = time_df, "Tukey_Celltypes" = cell_df, 
                     "Tukey_interaction" = df, "TukeySigInteraction" = sig_df)

write_xlsx(results.list, path = "ANOVA2TimeCellsProportion.xlsx")



```

Proportion test

```{r}
library(scProportionTest)

Idents(seu.t) <- "CelltypesMainNew"
prop_test <- sc_utils(seu.t)

# set up the comparison

# need to compare separately
prop_test <- permutation_test(
	prop_test, cluster_identity = "CelltypesMainNew",
	sample_1 = "30", sample_2 = "60",
	sample_identity = "Days_in_FD"
)


# make the plot
permutation_plot(prop_test)

pdf("prp30vs60allreps.pdf", width = 7, height = 4)
permutation_plot(prop_test) +  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),    # Adjust axis title size
  legend.text = element_text(size = 12),   # Adjust legend text size
  legend.title = element_text(size = 14))
dev.off()

prop_test <- permutation_test(
	prop_test, cluster_identity = "CelltypesMainNew",
	sample_1 = "60", sample_2 = "100",
	sample_identity = "Days_in_FD"
)
permutation_plot(prop_test)

pdf("prp60vs100allreps.pdf", width = 7, height = 4)
permutation_plot(prop_test) +  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),    # Adjust axis title size
  legend.text = element_text(size = 12),   # Adjust legend text size
  legend.title = element_text(size = 14))
dev.off()

prop_test <- permutation_test(
	prop_test, cluster_identity = "CelltypesMainNew",
	sample_1 = "100", sample_2 = "150",
	sample_identity = "Days_in_FD"
)
permutation_plot(prop_test)

pdf("prp100vs150allreps.pdf", width = 7, height = 4)
permutation_plot(prop_test) +  theme(axis.text.x = element_text(size = 12),
        axis.text.y = element_text(size = 12),
        axis.title = element_text(size = 14),    # Adjust axis title size
  legend.text = element_text(size = 12),   # Adjust legend text size
  legend.title = element_text(size = 14))
dev.off()



```


Proportion test by replicate

```{r}

unique(seu.t$Sample)

prop_test <- permutation_test(
	prop_test, cluster_identity = "CelltypesMainNew",
	sample_1 = "AIW_30_R1", sample_2 = "AIW_60_R1",
	sample_identity = "Sample"
)

permutation_plot(prop_test)


prop_test <- permutation_test(
	prop_test, cluster_identity = "CelltypesMainNew",
	sample_1 = "AIW_30_R2", sample_2 = "AIW_60_R1",
	sample_identity = "Sample"
)

permutation_plot(prop_test)

prop_test <- permutation_test(
	prop_test, cluster_identity = "CelltypesMainNew",
	sample_1 = "AIW_30_R3", sample_2 = "AIW_60_R1",
	sample_identity = "Sample"
)

permutation_plot(prop_test)


prop_test <- permutation_test(
	prop_test, cluster_identity = "CelltypesMainNew",
	sample_1 = "AIW_30_R4", sample_2 = "AIW_60_R1",
	sample_identity = "Sample"
)

permutation_plot(prop_test)

```



```{r}
# Load the data

# Load the tidyr package if not already loaded
# install.packages("tidyr")  # Uncomment and run if you haven't installed it

library(tidyr)
df <- as.data.frame(table(seu.t$CelltypesMainNew,seu.t$Days_in_FD,seu.t$Replicate))
colnames(df) <- c("Celltypes","DaysFD","Replicate","Freq")
# Pivot the data
# Pivot the data
fdata <- pivot_wider(df, names_from = DaysFD, values_from = Freq)
# Get unique cell types
unique_celltypes <- unique(df$Celltypes)

# Initialize an empty data frame to store results
results_df <- data.frame(Celltype = character(0), ChiSquare = numeric(0), P_Value = numeric(0))

# Loop through each unique cell type and perform chi-square test
for (celltype in unique_celltypes) {
  subset_df <- df[df$Celltypes == celltype, ]
  chisq_result <- chisq.test(subset_df$Freq)
  
  # Append the results to the data frame
  results_df <- rbind(results_df, data.frame(
    Celltype = celltype,
    ChiSquare = chisq_result$statistic,
    P_Value = chisq_result$p.value
  ))
}

# Print the results
print(results_df)

```

Multinomial Logistic Regression with Replicates

```{r}
library(nnet)

df <- as.data.frame(table(seu.t$CelltypesMainNew,seu.t$Days_in_FD,seu.t$Replicate))
colnames(df) <- c("Celltypes","DaysFD","Replicate","Freq")
# Assuming df is your data frame with cell type proportions for each sample
# Replicates are in the "Replicate" column
model <- multinom(Celltypes ~ DaysFD + Replicate, data = df)
summary(model)



```

