---
title: "May 10th and June 10th new hMOs"
output: html_notebook
---

Set up the workspace

```{r}

require("flowCore") #Used for reading the data
require("ggplot2")
require("ggridges") #visualization
require("stringr") #set of functions to manipulate strings type in order to have nice titles
require("rlist") #set of functions that allows to easily manipulate lists
require("reshape2") #visualization
require("flowStats") #Alignment functions
require("scales") #scale colour intensity for visualization
require("dplyr")
require("Seurat")
library(randomForest)
library('data.table')
# library("CelltypeR")
#source("/Users/rhalenathomas/GITHUB/CelltypeR/CelltypeR_functions/Preprocessing.R")


```

Preprocessing hMO datasets.  Live single cells gated in FlowJo

1. Set pathways

```{r}

input_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/GatingPlanExperiment/May10June10_livegating"

output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/FlowDataFiles/PresortingMay10June10/"

write_fcs_files <- TRUE #Set to true to write fcs files at each step (subsampled, transformed, aligned and scaled) - recommended 

#Create output folder if it's not already created
if(dir.exists(output_path)==FALSE){ #check
  dir.create(output_path) #create directory
}




```


2. Make flowset object

```{r}


flowset3 <- fsc_to_df(input_path = input_path)  
sampleNames(flowset3)

# for some when reading in both file sets together I get an error
# either wrong dimension or wrong type 

# rename samples 
sampleNames(flowset3) <- c("M10_A","M10_B","M10_C","M10_D",
                           "J10_A", "J10_B", "J10_C")
sampleNames(flowset3)


```


Check the data object


```{r}

flowset <- flowset3

plotdensity_flowset <- function(flowset){ ggplot(melt(lapply(as.list(flowset@frames),function(x){x=as.data.frame(x@exprs)})), aes(x=value,y=L1,fill=L1)) + geom_density_ridges(alpha=.4,verbose=FALSE) +facet_wrap(~variable)+theme_light()} 

plotdensity_flowset(flowset)

```
Note the ssc-h values are very different between May and June.  These value are not used in the clustering.


Need to make a combined function

```{r}

#### need to change this function to detect peaks or at least have the user add the column number
### check the density plots above to see which have peaks

transformed_flowset <- transform(flowset, transformList(colnames(flowset), biexp))

colnames(transformed_flowset)
normtr=gaussNorm(transformed_flowset,colnames(transformed_flowset)[c(1,2,3,9:length(colnames(transformed_flowset))-1)],max.lms = 2,peak.density.thr = 0.001)


expbe_norm2=normtr$flowset
normtr=gaussNorm(expbe_norm2,colnames(expbe_norm2)[c(4,5,7:8)],max.lms = 1,peak.density.thr = 0.05)#Detects and align 1 peak 
aligned_transformed_flowset=normtr$flowset
retrotransformed_flowset <- inversebiexponentialTransform(aligned_transformed_flowset) 


```


See the transformations

```{r}
plotdensity_flowset(flowset)
plotdensity_flowset(transformed_flowset)
plotdensity_flowset(aligned_transformed_flowset)


```
One sample On sample is far out of alignment in Cy5



Save the flowset dataset May10

```{r}

flowsetMayJune10 <- retrotransformed_flowset
flowset_to_csv(flowsetMayJune10, output_path)



```


Now create the Seurat object

```{r}

AB_vector <- c("AQP4", "CD24", "CD44","CD184","CD15","HepaCAM","CD29",
                         "CD56", "O4","CD140a","CD133","GLAST","CD71")

df <- read.csv(paste(output_path,"flowsetMayJune10.csv",sep=""))

#seu1 <- make_seu(df1, AB_vector)
#seu2 <- make_seu(df2, AB_vector)

seu <- make_seu(df, AB_vector)
# object is scaled in the function


```

Cluster


```{r}

# number of antibodies measured/ features 
n = 13
# list of antibodies - could also include FSC-W etc
#AB_vector 
i = 40
j = c(0.8,1.2)

AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")


seu <- ScaleData(seu, model.use = "negbinom")

seu <- RunPCA(seu, features = AB, npcs = n, approx = FALSE)
seu <- FindNeighbors(seu, dims = 1:13, k.param = i)
seu <- FindClusters(seu, resolution = 0.8)
seu <- RunUMAP(seu, dims = 1:12, n.neighbors = 46, min.dist = 0.4,
               spread = 1.5)
DimPlot(seu)



```
See dim plots

```{r}

saveRDS(seu, paste(output_path, "NewDataannaote.14122022.Rds"))

seu <- readRDS(paste(output_path,"NewDataannaote.14122022.Rds"))

DimPlot(seu, group.by = "RNA_snn_res.1.2")

```

Now annotate the cell types using:
1. correlation predictions
2. Visualization of heatmaps
3. RandomForest
4. Seurat transfer


```{r}

# correlation predictions

# run the correlation prediction function on the full dataset
reference_path <- "/Users/rhalenathomas/GITHUB/PhenoID_single_cell_flow_cytometry_analysis/old/correlation/ReferenceMatrix9celltypesOrdered.csv"
output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/FlowDataFiles/PresortingMay10June10/"

test_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/FlowDataFiles/PresortingMay10June10/flowsetMayJune10.csv"

ref <- read.csv("/Users/rhalenathomas/GITHUB/PhenoID_single_cell_flow_cytometry_analysis/old/correlation/ReferenceMatrix9celltypesOrdered.csv")

test <- read.csv(test_path)
#run the correlation function

cor <- find_correlation(df, # see above
                        ref, 
                             min_corr = 0.1, 
                             min_diff = 0.05)

write.csv(paste(output_path,"NewDataCorrelations.csv", sep = ""))

# this produces a df with the Frequencies of each cell type and a csv with the best and second best correlation coefficients


# add the correlation predictions to the meta data

seu <- AddMetaData(object=seu, metadata=cor$cell.label, col.name = 'cor.labels')
unique(seu$cor.labels)

saveRDS(seu, paste(output_path,"NewDataannaote.14122022.Rds"))

DimPlot(seu, group.by = "cor.labels")

# make a table of the predictions per cluster

t.lables <- as.data.frame(table(seu$RNA_snn_res.1.2,seu$cor.labels))
t.lables$Freq <- as.double(t.lables$Freq)
# take top 2 predictions in case of unknown being the top 
top.prediction <-as.data.frame(t.lables  %>% group_by(Var1)  %>% top_n(1, Freq))
topcor <- top.prediction

top.prediction <-as.data.frame(t.lables  %>% group_by(Var1)  %>% top_n(2, Freq))
topcor2 <- top.prediction

write.csv(topcor, paste(output_path,"corpredictions.csv"))


```



```{r}
# Random Forest predictions

# read in the saved model
rf <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/RFM_trained.11072022.Rds")

AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")
seu <- ScaleData(seu, model.use = "negbinom")
df <- transpose(as.data.frame(GetAssayData(seu,slot = 'scale.data')))
dim(df)
colnames(df) <- AB
rfm.pred <- predict(rf,df)
head(rfm.pred)
results.df <- as.data.frame(rfm.pred)
head(results.df)

# add the predictions to the seurat object
seu <- AddMetaData(object=seu, metadata= as.factor(results.df$rfm.pred), col.name = 'rf.pred')
unique(seu$rf.pred)
# save the RFM predictions
#saveRDS(seu, paste(output_path,"NewDataannaote.14122022.Rds"))

DimPlot(seu, group.by = "rf.pred")

# make a table of the predictions per cluster

t.lables <- as.data.frame(table(seu$RNA_snn_res.1.2,seu$rf.pred))
t.lables$Freq <- as.double(t.lables$Freq)
# take top 2 predictions in case of unknown being the top 
top.prediction <-as.data.frame(t.lables  %>% group_by(Var1)  %>% top_n(1, Freq))
top1rf <- top.prediction
write.csv(top1rf, paste(output_path,"RFpredictions1.csv"))



```

Why doesn't this work - I don't know

```{r}

# seurat transfer anchor predictions
# using the 9000 cells reference used for the full dataset annotation

seu.r<- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/Seu9000annot.08072021.RDS")


outpath <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/"
seu.r <- readRDS(paste(outpath,"All9MOannaote.12072022.Rds"))
table(seu.r$cell.types)
table(seu.r$subgroups)
table(seu.r$most.subgroups)
DefaultAssay(seu.r) <- 'RNA'

Idents(seu.r) <- 'cell.types'
seu.r <- subset(seu.r, downsample = 5000)
table(seu.r$cell.types)
#seu.r <- ScaleData(seu.r, model.use = "negbinom")
#seu.r <- NormalizeData(seu.r)
#seu.r <- RunPCA(seu.r, features = AB)

#seu <- NormalizeData(seu)
#seu <- ScaleData(seu)
#seu.r <- ScaleData(seu.r)

AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

# we don't have a find variable features input, I have set all the features as th input
seu.r <- NormalizeData(seu.r)
seu <- NormalizeData(seu)
seu.r <- ScaleData(seu.r, model.use = 'negbinom')
seu <- ScaleData(seu, model.use = 'negbinom')
seu.r <- ScaleData(seu.r)
seu <- ScaleData(seu)

anchors <- FindTransferAnchors(reference = seu.r, query = seu, features = AB, dim= 1:12, npcs = 10, reduction = 'rpca') 

# retains 6 anchors only  but so many more are retained in the other data > 1500
# must set
predictions <- TransferData(anchorset = anchors, refdata = seu.r$cell.types, dims = 1:10, k.weight = 9)

seu <- AddMetaData(seu, predictions$prediction.score.max, col.name = "trans.pred")
DimPlot(seu, group.by = 'trans.pred')


```


```{r}

RidgePlot(seu.r, features = "CD24", slot = 'counts', group.by = 'orig.ident', log = TRUE)

RidgePlot(seu.r, features = "CD24", slot = 'scale.data', group.by = 'orig.ident', log = FALSE)

RidgePlot(seu, features = "CD24", slot = 'counts', group.by = 'orig.ident', log = TRUE)

RidgePlot(seu, features = "CD24", slot = 'scale.data', group.by = 'orig.ident', log = FALSE)

```


Visualization 


```{r}

DoHeatmap(seu, group.by = 'RNA_snn_res.1.2', features = AB, size= 5,slot = "scale.data",
        angle = 90,label = TRUE,
        hjust = 0, group.bar.height = 0.05, disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 13)) 

```


Feature plots

```{r}

Idents(seu) <- 'RNA_snn_res.1.2'
FeaturePlot(seu, features = AB, slot = 'scale.data',min.cutoff = 'q3', max.cutoff ='q97', ncol = 5, label = TRUE)

for(A in AB){
  print(FeaturePlot(seu, features = A, slot = 'scale.data',min.cutoff = 'q3', max.cutoff ='q97', ncol = 1, label = TRUE))
}

```


Add cell type labels using subtypes from 9000 labels (used to train RF)


```{r}

Idents(seu) <- "RNA_snn_res.1.2"

# labels with the 9000 subset subgroup labels
cluster.ids <- c("mixed","astro1a","astro1b","astro2","RG1a",   # 0-4
                 "Neurons2","RG2","Neurons1a","RG1c","astro1c",  #5-9
                 "Neurons2a","stem-like", "Mixed","RG1b", "astro1d", #10-14
                 "Neurons3a", "OPC", "astro2b","stem-like","astro1e",  #15-19
                 "Neurons3b", "endothelial","astro1f", "Neurons2b","Neurons3b", #20-24
                 "Neurons1b","Neurons3c","RG3","NPC","NPCb", #25-29
                 "astro3","astro-m","epithelial","oligodendrocytes", "astro-m",
                 "RG/astro2")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$labels1 <- Idents(seu)

Idents(seu) <- "RNA_snn_res.1.2"
# combine same 
cluster.ids <- c("Mixed","Astrocytes 1","Radial Glia 3","Astrocytes 2","Radial Glia 1",   # 0-4
                 "Neurons 2","Radial Glia 2","Neurons 1","Radial Glia 2","Astrocytes 1",  #5-9
                 "Neurons 2","Stem-like", "Mixed","Radial Glia 1", "Astrocytes 1", #10-14
                 "Neurons 3", "OPC", "Astrocytes 2","Stem-like","Astrocytes 1",  #15-19
                 "Neurons 3", "Endothelial","Astrocytes 1", "Neurons 2","Neurons 3", #20-24
                 "Neurons 1","Neurons 3","Radial Glia 3","NPC","NPC", #25-29
                 "Astrocytes 3","Astrocytes mature","Epithelial","Oligodendrocytes", "Astrocytes mature",
                 "Glia")

# combine same 
names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$subtypes <- Idents(seu)
DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'subtypes', repel = TRUE) +
  ggplot2::theme(legend.position = "none")

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'subtypes', repel = TRUE)


```

```{r}

DoHeatmap(seu, group.by = 'subtypes', features = AB, size= 5,slot = "scale.data",
        angle = 90,label = TRUE,
        hjust = 0, group.bar.height = 0.05, disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 13)) 

```

Labels with groups matching for proportions

```{r}

Idents(seu) <- "RNA_snn_res.1.2"

# main group labels for proportions

cluster.ids <- c("Mixed","Astrocytes 1","Radial Glia 1","Astrocytes 2","Radial Glia 1",   # 0-4
                 "Neurons 2","Radial Glia 2","Neurons 1","Radial Glia 2","Astrocytes 1",  #5-9
                 "Neurons 2","Stem-like", "Mixed","Radial Glia 1", "Astrocytes 1", #10-14
                 "Neurons 1", "OPC", "Astrocytes 2","Stem-like","Astrocytes 1",  #15-19
                 "Neurons 1", "Endothelial","Astrocytes 1", "Neurons 2","Neurons 2", #20-24
                 "Neurons 1","Neurons 2","Radial Glia 1","NPC","NPC", #25-29
                 "Astrocytes 1","Astrocytes 2","Epithelial","Oligodendrocytes", "Astrocytes 2",
                 "Glia")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$Cell_types <- Idents(seu)

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'Cell_types', repel = TRUE) +
  ggplot2::theme(legend.position = "none")

DimPlot(seu, reduction = "umap", label = TRUE, group.by = 'Cell_types', repel = TRUE)


DimPlot(seu, group.by = 'RNA_snn_res.1.2', label = TRUE)

```


```{r}

saveRDS(seu, paste(output_path, "NewDataannaote.14122022.Rds"))


```

Heatmaps

```{r}
DoHeatmap(seu, group.by = 'Cell_types', features = AB, size= 5,slot = "scale.data",
        angle = 90,label = TRUE,
        hjust = 0, group.bar.height = 0.05, disp.max = 2, disp.min = -1.5) + scale_fill_gradientn(colors = c("#154c79", "#eeeee4", "#e28743")) + 
  theme(axis.text.y = element_text(size = 13)) 

```


Cell Types for figures

```{r}

Idents(seu) <- "RNA_snn_res.1.2"

cluster.ids <- c("Mixed","Astrocytes 1","Radial Glia 1","Astrocytes 2","Radial Glia 1",   # 0-4
                 "Neurons 2","Radial Glia 2","Neurons 1","Radial Glia 2","Astrocytes 1",  #5-9
                 "Neurons 2","Epithelial", "Neurons 2","Radial Glia 1", "Astrocytes 1", #10-14
                 "Neurons 1", "OPC", "Astrocytes 2","Stem cell like","Astrocytes 1",  #15-19
                 "Neurons 1", "Endothelial","Astrocytes 1", "Neurons 2","Neurons 2", #20-24
                 "Neurons 1","Neurons 2","Radial Glia 1","NPC","NPC", #25-29
                 "Astrocytes 1","Astrocytes 2","Epithelial","Oligodendrocytes", "Astrocytes 2",
                 "Glia")

names(cluster.ids) <- levels(seu)
seu <- RenameIdents(seu, cluster.ids)
seu$Cell_types <- Idents(seu)


# change the order of the cell types on the legend of the umap
cell.type.order <- c("Astrocytes 1", "Astrocytes 2","Radial Glia 1","Radial Glia 2",
                     "Epithelial","Endothelial","NPC","Neurons 1","Neurons 2",
                     "Oligodendrocytes","OPC","Stem cell like","Glia","Mixed")
cell.type.order <- rev(cell.type.order)

# colour order to match cell type order
clust.colours <- c("chocolate1","orange","lightsalmon", "pink",
                   "steelblue3","deepskyblue","plum3","purple","orchid2",
                   "seagreen3","green","tomato3","burlywood3","grey90","lemonchiffon3","yellow")

Idents <- 'Cell_types'
DimPlot(seu, order = cell.type.order, cols = clust.colours, shuffle = TRUE, raster=FALSE,pt.size = 0.1, label = FALSE, label.size = 6) +
  theme(legend.text = element_text(size=16), axis.title.y = element_text(size=16), 
        axis.title.x = element_text(size=16), axis.text.y = element_text(size =16),
        axis.text.x = element_text(size =16))





```







