---
title: "May 10th and June 10th new hMOs"
output: html_notebook
---

Set up the workspace

```{r}

require("flowCore") #Used for reading the data
require("ggplot2")
require("ggridges") #visualization
require("stringr") #set of functions to manipulate strings type in order to have nice titles
require("rlist") #set of functions that allows to easily manipulate lists
require("reshape2") #visualization
require("flowStats") #Alignment functions
require("scales") #scale colour intensity for visualization
require("dplyr")
require("Seurat")
#library(randomForest)
library('data.table')
# library("CelltypeR")
#source("/Users/rhalenathomas/GITHUB/CelltypeR/CelltypeR_functions/Preprocessing.R")


```

Preprocessing hMO datasets.  Live single cells gated in FlowJo

1. Set pathways

```{r}

input_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/GatingPlanExperiment/May10June10_livegating/"

output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/FlowDataFiles/PresortingMay10June10/"

write_fcs_files <- TRUE #Set to true to write fcs files at each step (subsampled, transformed, aligned and scaled) - recommended 

#Create output folder if it's not already created
if(dir.exists(output_path)==FALSE){ #check
  dir.create(output_path) #create directory
}




```


2. Make flowset object

```{r}

flowset <- read.flowSet(path=input_path) #Create a flowset object

# see if each dataset can be read in separately or if the JuneGating has a problem. 
# Somehow May and June might not be 
flowset1 <- read.flowSet(path="/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/GatingPlanExperiment/livegatingMay10/")

flowset2 <- read.flowSet(path="/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/GatingPlanExperiment/livegatingJune10/")

flowset3 <- fsc_to_df("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/GatingPlanExperiment/May10June10_livegating/")  
sampleNames(flowset1)
sampleNames(flowset2)
# for some when reading in both file sets together I get an error
# either wrong dimension or wrong type 

# rename samples 
sampleNames(flowset1) <- c("B1_M10_PBS","B1_M10_TRIS","B2_M10_PBS","B2_M10_TRIS")
sampleNames(flowset1)

sampleNames(flowset2) <- c("B1_J10_PBS_R1","B1_J10_PBS_R2","B1_J10_PBS_R3")
sampleNames(flowset2) 

```


Check the data object


```{r}

flowset <- flowset1
flowset <- flowset2
plotdensity_flowset <- function(flowset){ ggplot(melt(lapply(as.list(flowset@frames),function(x){x=as.data.frame(x@exprs)})), aes(x=value,y=L1,fill=L1)) + geom_density_ridges(alpha=.4,verbose=FALSE) +facet_wrap(~variable)+theme_light()} 

plotdensity_flowset(flowset)

```

try to fix the align function

```{r}
##### these are for flowset1
#### need to change this function to detect peaks or at least have the user add the column number
colnames(transformed_flowset)
normtr=gaussNorm(transformed_flowset,colnames(transformed_flowset)[c(1,3,9:length(colnames(transformed_flowset))-1)],max.lms = 2,peak.density.thr = 0.001)


expbe_norm2=normtr$flowset
normtr=gaussNorm(expbe_norm2,colnames(expbe_norm2)[c(4,7:8)],max.lms = 1,peak.density.thr = 0.05)#Detects and align 1 peak 
aligned_transformed_flowset=normtr$flowset
retrotransformed_flowset <- inversebiexponentialTransform(aligned_transformed_flowset) 

```

Alignment in flowset2 (June10)

```{r}

# same conditions work as in May10th
#### need to change this function to detect peaks or at least have the user add the column number
colnames(transformed_flowset)
normtr=gaussNorm(transformed_flowset,colnames(transformed_flowset)[c(1,3,9:length(colnames(transformed_flowset))-1)],max.lms = 2,peak.density.thr = 0.001)


expbe_norm2=normtr$flowset
normtr=gaussNorm(expbe_norm2,colnames(expbe_norm2)[c(4,7:8)],max.lms = 1,peak.density.thr = 0.05)#Detects and align 1 peak 
aligned_transformed_flowset=normtr$flowset
retrotransformed_flowset <- inversebiexponentialTransform(aligned_transformed_flowset) 
```



See the transformations
```{r}
plotdensity_flowset(flowset)
plotdensity_flowset(transformed_flowset)
plotdensity_flowset(aligned_transformed_flowset)


```


Save the flowset dataset May10

```{r}

flowsetMay10 <- retrotransformed_flowset
flowset_to_csv(flowsetMay10)


```

Save flowset from June 10

```{r}
flowsetJune10 <- retrotransformed_flowset
flowset_to_csv(flowsetJune10)

```

Now create the Seurat object

```{r}

AB_vector <- c("AQP4", "CD24", "CD44","CD184","CD15","HepaCAM","CD29",
                         "CD56", "O4","CD140a","CD133","GLAST","CD71")

df1 <- read.csv(paste(output_path,"flowsetMay10.csv",sep=""))
df2 <- read.csv(paste(output_path,"flowsetJune10.csv",sep=""))
dim(df1)
dim(df2)

# The csv doesn't have both datasets May and June
df <- rbind(df1,df2)
dim(df)
df$cell <- rownames(df)
write.csv(df, paste(output_path, "FlowsetMayJune.csv"))

#seu1 <- make_seu(df1, AB_vector)
#seu2 <- make_seu(df2, AB_vector)

seu <- make_seu(df, AB_vector)
# object is scaled in the function


```

Cluster


```{r}


# number of antibodies measured/ features 
n = 13
# list of antibodies - could also include FSC-W etc
#AB_vector 
i = 60
j = c(1.2)

AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

seu <- ScaleData(seu, model.use = 'binomial')

seu <- RunPCA(seu, features = AB, npcs = 12, approx = FALSE)
seu <- FindNeighbors(seu, dims = 1:12, k.param = i)
seu <- FindClusters(seu, resolution = j)
seu <- RunUMAP(seu, dims = 1:12, n.neighbors = i, min.dist = 0.1,
               spread = 1.3)






```
See dim plots

```{r}

seu <- readRDS(paste(output_path,"NewDataannaote.14122022.Rds"))
DimPlot(seu, group.by = 'RNA_snn_res.1.2', label = TRUE)


```

Now annotate the cell types using:
1. correlation predictions
2. Visualization of heatmaps
3. RandomForest
4. Seurat transfer


```{r}

# correlation predictions

# run the correlation prediction function on the full dataset
reference_path <- "/Users/rhalenathomas/GITHUB/PhenoID_single_cell_flow_cytometry_analysis/old/correlation/ReferenceMatrix9celltypesOrdered.csv"
output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/FlowDataFiles/PresortingMay10June10/"

test_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/FlowDataFiles/PresortingMay10June10//FlowsetMayJune.csv"


ref <- read.csv("/Users/rhalenathomas/GITHUB/PhenoID_single_cell_flow_cytometry_analysis/old/correlation/ReferenceMatrix9celltypesOrdered.csv")

#run the correlation function

cor <- find_correlation(df, # see above
                        ref, 
                             min_corr = 0.1, 
                             min_diff = 0.05)

write.csv(paste(output_path,"NewDataCorrelations.csv", sep = ""))

# this produces a df with the Frequencies of each cell type and a csv with the best and second best correlation coefficients





# add the correlation predictions to the meta data

seu <- AddMetaData(object=seu, metadata=cor$cell.label, col.name = 'cor.labels')
unique(seu$cor.labels)

saveRDS(seu, paste(output_path,"NewDataannaote.14122022.Rds"))

DimPlot(seu, group.by = "cor.labels")

# make a table of the predictions per cluster

t.lables <- as.data.frame(table(seu$RNA_snn_res.1.2,seu$cor.labels))
t.lables$Freq <- as.double(t.lables$Freq)
# take top 2 predictions in case of unknown being the top 
top.prediction <-as.data.frame(t.lables  %>% group_by(Var1)  %>% top_n(2, Freq))
top2cor <- top.prediction



```



```{r}
# Random Forest predictions

# read in the saved model
rf <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/RFM_trained.11072022.Rds")

AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

df <- transpose(as.data.frame(GetAssayData(seu,slot = 'scale.data')))
dim(df)
colnames(df) <- AB
rfm.pred <- predict(rf,df)
head(rfm.pred)
results.df <- as.data.frame(rfm.pred)
head(results.df)

# add the predictions to the seurat object
seu <- AddMetaData(object=seu, metadata= as.factor(results.df$rfm.pred), col.name = 'rf.pred')
unique(seu$rf.pred)
# save the RFM predictions
#saveRDS(seu, paste(output_path,"NewDataannaote.14122022.Rds"))

DimPlot(seu, group.by = "rf.pred")

# make a table of the predictions per cluster

t.lables <- as.data.frame(table(seu$RNA_snn_res.1.2,seu$rf.pred))
t.lables$Freq <- as.double(t.lables$Freq)
# take top 2 predictions in case of unknown being the top 
top.prediction <-as.data.frame(t.lables  %>% group_by(Var1)  %>% top_n(2, Freq))
top2rf <- top.prediction
write.csv(top2rf, paste(output_path,"RFpredictions.csv"))

```



```{r}

# seurat transfer anchor predictions
# using the 9000 cells reference used for the full dataset annotation

seu.q <- readRDS(paste(output_path,"NewDataannaote.14122022.Rds"))
DefaultAssay(seu) <- 'RNA'

seu.r<- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/Seu9000annot.08072021.RDS")
DefaultAssay(seu.r) <- 'RNA'

Idents(seu.r) <- 'subgroups'
seu.r <- subset(seu.r, downsample = 500)
table(seu.r$subgroups)
AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")

seu.r <- ScaleData(seu.r)
seu <- ScaleData(seu.q)
# we don't have a find variable features input, I have set all the features as th input
anchors <- FindTransferAnchors(reference = seu.r, query = seu, features = AB, dim= 1:10, npcs = 10, max.features = 13, reference.reduction = 'pca') 

# retains 6 anchors only  but so many more are retained in the other data > 1500
# must set
predictions <- TransferData(anchorset = anchors, refdata = seu.r$subgroups, dims = 1:10, k.weight = 4)


# errors might resolve afer restarting
predictions <- TransferData(anchorset = anchors, refdata = seu.r$subgroups, dims = 1:10, k.weight = 6)

seu <- AddMetaData(seu, predictions$prediction.score.max, col.name = "trans.pred")

DimPlot(seu, group.by = 'trans.pred')

RidgePlot(seu.q, features = "CD24", slot = 'counts', group.by = 'orig.ident', log = TRUE)

RidgePlot(seu.q, features = "CD24", slot = 'scale.data', group.by = 'orig.ident', log = FALSE)

RidgePlot(seu.r, features = "CD24", slot = 'counts', group.by = 'orig.ident', log = TRUE)

RidgePlot(seu.r, features = "CD24", slot = 'scale.data', group.by = 'orig.ident', log = FALSE)

RidgePlot(seu, features = "CD24", slot = 'counts', group.by = 'orig.ident', log = TRUE)

RidgePlot(seu, features = "CD24", slot = 'scale.data', group.by = 'orig.ident', log = FALSE)

```

Visualization 

```{r}

DoHeatmap(seu, features = AB)


```


Feature plots

```{r}

FeaturePlot(seu, features = AB, slot = 'scale.data',min.cutoff = 'q3', max.cutoff ='q97', ncol = 5, label = TRUE)

for(A in AB){
  print(FeaturePlot(seu, features = A, slot = 'scale.data',min.cutoff = 'q3', max.cutoff ='q97', ncol = 1, label = TRUE))
}

```





