---
title: "CelltypeR workflow"
output: html_notebook
---

1. Data preprocessing
a.	Read FlowJo files into R.
b.	Create a data frame with intensity measurements for each marker for all samples within the experiment to be analyzed.  
c.	Harmonize data 
d.	Explore clusters

```{r}
# load necessary libraries 
library(Seurat)
library(dplyr) 
library(ggplot2)
# library(CelltypeR)

```


Read in the flow data
This data should be the gated live cells.  
All samples need to be in one folder.


```{r}

input_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/FlowDataFiles/9MBO"
output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/testingLibrary"

# 1.a Read in FlowJo Files 

# choose to downsample to 9000 where possible
# check this - I don't think it worked
flowset <- fsc_to_fs(input_path, downsample = 9000)
# down sample can be a number, 'none' or 'min'

# look at file names and rename with shorter sample names

sampleNames(flowset)
sampleNames(flowset) <- sampleNames(flowset) <- c("3450_0306","AIW002_0306","AJG001C_0306","3450_0317A","AIW002_0317A","AJG001C_0317A","3450_0317B","AIW002_0317B","AJG001C_0317B")
sampleNames(flowset)


# if we want to save the flowset object now we can 
# this would be read back in with flowset 
# 



```


Harmonize data to remove batch or technical variation

This requires us to look and see where there are two peaks to align. We need to visualize the peaks of the transformed data before aligning.

```{r}

# we can decided what level of processing to choose with the argument 'processing'
# biexp only applies a biexponential transformation
# align applies biexp transform and then aligns peaks
# retro (default), transforms, aligns and then reverse transforms
flowset_biexp <- harmonize(flowset, processing = 'biexp')
# we can visualize the peaks to see where there are two peaks for alignment

# we need to enter the column index for which peaks to align, the alignment for one or two peaks is not the same. 
#plotdensity_flowset(flowset)
plotdensity_flowset(flowset_biexp) # to see the peaks

flowset_align <- harmonize(flowset, processing = 'align', 
                           two_peaks = c(7:20),
                       one_peak = c(1:6,21), threshold = 0.01)

flowset_retro <- harmonize(flowset, processing = 'retro', 
                           two_peaks = c(7:20),
                       one_peak = c(1:6,21), threshold = 0.01)

df <- flowset_to_csv(flowset_retro)




```

Now we have made all the different processing of the fsc files.  We can visualize the intensity in cell density plots to see the alignment

```{r}
#plotdensity_flowset(flowset)
plotdensity_flowset(flowset_biexp)
plotdensity_flowset(flowset_align)
#plotdensity_flowset(flowset_retro)


```

Now we will make test out clustering.
For Seurat clustering and Phenograph we will make the Seurat object
Flowsome takes in the dataframe directly.


```{r}

# the function make_seu will take in the df of expression and Antibody/Marker list as a vector and create a seurat object. Values are scaled. Marker expression will be in the "RNA" slot. PCA is calculated using AB vector as the features 

# make sure to always keep the same antibody order or your labels will not be correct


# antibody features in order to appear on the plots
AB <- c("CD24","CD56","CD29","CD15","CD184","CD133","CD71","CD44","GLAST","AQP4","HepaCAM", "CD140a","O4")
seu <- make_seu(df, AB_vector = AB)


```
Save dataframe and seurat object for later

```{r}

output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/testingLibrary/"

# to save the df for later
# write.csv(df, "pathway/filename.csv")
write.csv(df, paste(output_path,"df9000Feb15.csv", sep = ""))

# save the seurat object
saveRDS(seu, paste(output_path,"seu9000Feb15.RDS", sep = ""))



```


Read in the csv of the flow files processed and the seurat object

```{r}
df.input <- read.csv("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/testingLibrary/df9000Feb15.csv")

seu <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/testingLibrary/seu9000Feb15.RDS")


```

Test Flowsom

```{r}

output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/testingLibrary/exp_clusters/flow/"

flowsom.test <- flowsom(input = seu, #seurat
                    df_input = df.input, #the processed df2 file before being converted to seurat
                    flow_k = c(3,5),
                    run.stats = TRUE,
                    run.plot = FALSE,
                    save_to = output_path)

```



Explore cluster parameters

```{r}
df2 <- df.input %>% dplyr::select(AB) # need to add this line into the main explore param function 

```

Phenograph

```{r}

output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/testingLibrary/exp_clusters/pheno/"

pheno.test <- phenograph(input = seu, df_input = df.input,
                         pheno_lou_kn = c(20,40),
                         run.stats = TRUE,
                         run.plot = TRUE,
                         save_to = output_path)


```


Test functions Louvain

```{r}

output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/testingLibrary/exp_clusters/seu/"

lou.test <- louvain(input = seu, #seu object
                    df_input = df.input,
                    pheno_lou_kn = c(20,40),
                    resolutions = c(0.2,0.5),
                    run.plot = FALSE, #option to save the graphs  
                    run.stats = FALSE, #option to save stats list  
                    save_to #only required when save is TRUE
)


```





Annotate clusters
1. Visualization for manual annotation. - output by clustering function
2. CAM (Correlation assignment model) - requires reference matrix
3. RFM (Random Forest Model) - requires annotated matching flow dataset
4. Seurat label transfer - requires annotated matching flow data in a seurat object



```{r}

# run clustering with only desired conditions - use function - TEMP run directly 

seu <- RunPCA(seu, features = AB)
seu <- FindNeighbors(seu, dims = 1:12, k.param = 60)
# must take one less than the number of antibodies 

seu <- FindClusters(seu, resolution = 0.8)
seu <- RunUMAP(seu, dims = 1:12, n.neighbors = 46, min.dist = 0.4,
               spread = 1.5)
DimPlot(seu)


```

```{r}
# save with graph
saveRDS(seu,"/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/testingLibrary/seu9000Feb24.RDS")

```



Visualize expression on UMAP and with heat maps

```{r}

# this will let us see one at at time
for (i in AB) {
  print(FeaturePlot(seu, features = i, min.cutoff = 'q1', max.cutoff = 'q95', label = TRUE))
}


```



```{r}

output_path <- "/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/testingLibrary/"
saveRDS(seu, paste(output_path,"cluster9000.RDS"))

```

Predict cell annotations with CAM (Corralations assignment method)

```{r}

reference_path <- "/Users/rhalenathomas/GITHUB/CelltypeR/Data/ReferenceMatrix9celltypesOrdered.csv"



test_data <- read.csv("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/testingLibrary/df9000Feb15.csv") 
reference_data <- read.csv(reference_path)

cor <- find_correlation(test_data, 
                             reference_data, 
                             min_corr = 0.3, 
                             min_diff = 0.05)

class(cor)
# creates a dataframe with cor1 cor2 and predicted cell type label

```

Visualize the CAM results

```{r}

plot_corr(cor)


```

Apply correlation predictions to clusters and output a vector for annotation functions

```{r}


seu <- AddMetaData(object=seu, metadata=cor$cell.label, col.name = 'cor.labels')
# see the labels added
unique(seu$cor.labels)

seu.cluster = seu$RNA_snn_res.0.8
seu.labels = seu$cor.labels

# plot the cluster predictions
plot_lab_clust(seu, seu$RNA_snn_res.0.8, seu$cor.labels)



```

Run get annotations function to return a vector of annotation in the order of the clusters.

```{r}


cor.ann <- get_annotation(seu, seu.cluster = seu$RNA_snn_res.0.8, seu.label = seu$cor.labels, 
               top_n = 3, ignore_unknown = TRUE, Label = "CAM")
cor.ann




```


Use a trained Random Forest model to predict cell types. 
Training of the Random Forest model with an annotated data set is below.

```{r}

# you must have a saved trained model from a data object annotated from the same markers

rf <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/RFM_trained.11072022.Rds")


rfm.pred <- RFM_predict(seu, rf)
head(rfm.pred)

# add the predictions into the seurat object

seu <- AddMetaData(object=seu, metadata=rfm.pred$`predict(rf, df)`, col.name = 'rfm.labels')

# check that the data is added 
table(seu$rfm.labels)


```

Get the annotation by cluster for the RFM

```{r}

rfm.ann <- get_annotation(seu, seu$RNA_snn_res.0.8,seu$rfm.labels, 
               top_n = 3, filter_out = c("unknown","Unknown","Mixed"), Label = "RFM")
rfm.ann


```

Plot RFM predictions

```{r}


plot_lab_clust(seu, seu.cluster = seu$RNA_snn_res.0.8, seu.labels = seu$rfm.labels, filter_out = c("unknown","Unknown","Mixed"))


```



Predicting cell types with Seurat label transfer using anchors



```{r}

# takes in a seurat object with the labels added 
# makes a dataframe with the count of predicted labels for each cluster
# input seurat object with the predicted labels in the meta data
# input the clusters meta data slot to be labels
# input the meta data slot with the labels (correlation, random forest, seurat predicted)

#need reference data object with labels
seu.r<- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/PaperFigures/Seu9000annot.08072021.RDS")


# the output is a seurat object with the predicted annotations

seu <- seurat_predict(seu, seu.r, ref_id = 'subgroups', down.sample = 500, markers = AB)




```

```{r}

# plot the seurat anchor predictions
# get the annotation table for the seurat anchor predictions 

plot_lab_clust(seu, seu$RNA_snn_res.0.8, seu$seu.pred)

# to not filter anything use c()
seu.ann <- get_annotation(seu, seu$RNA_snn_res.0.8,seu$seu.pred, 
               top_n = 3, filter_out = c(), Label = "Seurat")
seu.ann


```

Get a consensus of cluster annotations, Add the annotations to the seurat object



```{r}

# manual annotations
# quick cheat just using the seurat labels to get a sting of annotations

temp <- paste("'",as.character(seu.ann$Seurat), "'", collapse = ", ", sep = "" )
my.ann <- c('Unknown', 'Radial Glia 1', 'Unknown', 'Mixed', 'Neurons 1', 'Radial Glia 3', 'Epithelial', 'Astrocytes mature', 'Endothelial', 'Astrocytes 1', 'Radial Glia 2', 'Neurons 1', 'NPC', 'Neurons 3', 'Epithelial', 'Neurons 1', 'Epithelial', 'Oligodendrocytes', 'Stem-like 1', 'Epithelial', 'NPC', 'Astrocytes mature', 'Oligodendrocytes', 'Stem-like 1')
cluster.n <- c(0:23)
man.ann <- data.frame(cluster.n, my.ann)
colnames(man.ann) <- c("Cluster","manual")

# needs to be factors 
man.ann <- data.frame(lapply(man.ann, factor))

dim(man.ann)
dim(cor.ann)
dim(rfm.ann)
dim(seu.ann)
# function to take in dataframes of different predictions and a vector of manual annotation
# only one is required

# the annotation function takes in a list of the annotation dataframes

ann.list <- list(man.ann,cor.ann,rfm.ann,seu.ann)

# annotate the seurat object

seu <- cluster_annotate(seu, ann.list, 
                        annotation_name ="CellType3", 
                        to_label = "RNA_snn_res.0.8")


```


```{r}
DimPlot(seu)

```

Save an annotated data object

```{r}


```



Compare groups

```{r}

seu <- readRDS("/Users/rhalenathomas/Documents/Data/FlowCytometry/PhenoID/Analysis/testingLibrary/seu9000Feb15.RDS")


```










Train Random Forest model
Requires a labelled seurat object
More cells will give higher accuracy but increase computation time.
Run in HPC with lots of cells


```{r}

rf <- RFM_train(seurate_object = seu, 
                             AB_list = AB, annotations = seu$CellType3,
                      split = c(0.8,0.2),
                      downsample = 20000,
                      seed = 222,
                      mytry = c(1:10),
                      maxnodes = c(12: 25),
                      trees = c(250, 500, 1000,2000),
                      start_node = 15)

save(rf, output_path,"trainedRFMFeb14.Rds")

```


